---
title: "MT_03_Models"
author: "Fabian Gienke"
format: 
  html:
    toc: true           # Inhaltsverzeichnis am Rand
    code-fold: false     
    code-summary: "Code anzeigen"
    theme: cosmo
execute:
  cache: true   
  results: 'hide'
  message: false        
editor_options: 
  chunk_output_type: console
---

## STM and GT

This part of the code runs all models and pre-models like searchK. This includes STM specific data processing, STM Model 1 (no covariates), STM Model 2 (with covariates) for both TS and NYT. Also data is aggregated weekly to match GT. The GT data is added and processed to match the structure. The only Input Data needed is the TS & NYT Output from MT_02_Preprocessing for STM and the Keyword_List.xlsx for GT.

Packages:

```{r}
library(stm)
library(tidyverse)
library(quanteda)
library(textstem)
library(stopwords)
library(readxl)
library(dplyr)
library(purrr)
```

## Load Data

The data input consists of TS_Input and NYT_Input.

The file path needs to be changed accordingly.

```{r}
#| label: Load Data
# Input data
TS_Input <- read.csv('/TS_Input.csv', row.names = 1)
NYT_Input <- read.csv('/NYT_Input.csv', row.names = 1)

```

## STM Preprocessing for TS

The quanteda packages is applied for STM processing and data cleaning as discussed in Section 3 Methods (Masterthesis). This includes insuring right data types, removing punctuation, numbers, symbols, seperators. Also lower casing and stopword are removed. Then building a DFM (Data Frequency Matrix) for the STM Input. Covariates are added for the second model for robustness checks and include date, country_iso (as identified by country finder) and country_n (numbers of countries mentioned in each document).

```{r}
#| label: Preprocessing TS for STM
# Preprocessing: Lemmatization to reduce words to basic form
TS_Input$text_lemma <- lemmatize_strings(TS_Input$text)

# Preprocessing: quanteda preprocessing Corpus (more control than basic stm preprocessing functions)
TS_Input$doc_id <- as.character(1:nrow(TS_Input)) 
corpus_TS <- corpus(TS_Input, text_field = "text_lemma", docid_field = "doc_id")

# Preprocessing: Tokenization to remoce punctuation, numbers
tokens_TS <- quanteda::tokens(
  corpus_TS,
  remove_punct = TRUE,
  remove_numbers = TRUE,
  remove_symbols = TRUE,
  remove_separators = TRUE)

tokens_TS <- tokens_tolower(tokens_TS) # Lowercase
tokens_TS <- tokens_remove(tokens_TS, stopwords(language = "en", source = "snowball")) # Remove stopwords with smart (n=571) list or snowball base list (178)

dfm_TS <- dfm(tokens_TS) # Preprocessing: DFM 
dfm_TS <- dfm_trim(dfm_TS, min_termfreq = 10) # Trim rare words <10

# Convert quanteda objects to STM format (STM inputs)
stm_input_TS <- convert(dfm_TS, to = "stm")
documents_TS <- stm_input_TS$documents
vocab_TS <- stm_input_TS$vocab
meta_TS <- TS_Input[match(docnames(dfm_TS), TS_Input$doc_id), ]





# TS // define metadata / covariates for the STM Input
meta_TS$date_numeric <- as.numeric(as.Date(meta_TS$date))
meta_TS$country_iso <- ifelse(is.na(meta_TS$country_iso), "none", as.character(meta_TS$country_iso))
meta_TS$country_n <- ifelse(is.na(meta_TS$country_n), "none", as.character(meta_TS$country_n))

meta_TS$country_n <- as.factor(meta_TS$country_n)
meta_TS$country_iso <- as.factor(meta_TS$country_iso)
```

## STM Preprocessing for NYT

The NYT processing follows the exact steps of the TS processing.

```{r}
#| label: Preprocessing NYT for STM
# PART 2 for NYT (identical to TS processing)
set.seed(2026)
# Lemmatization to reduce words to basic form
NYT_Input$text_lemma <- lemmatize_strings(NYT_Input$text)

# quanteda preprocessing Corpus (more control than basic stm preprocessing functions)
NYT_Input$doc_id <- as.character(1:nrow(NYT_Input))
corpus_NYT <- corpus(NYT_Input, text_field = "text_lemma", docid_field = "doc_id")
# Tokenization to remoce punctuation, numbers
tokens_NYT <- quanteda::tokens(
  corpus_NYT,
  remove_punct = TRUE,
  remove_numbers = TRUE,
  remove_symbols = TRUE,
  remove_separators = TRUE)

tokens_NYT <- tokens_tolower(tokens_NYT) # Lowercase
tokens_NYT <- tokens_remove(tokens_NYT, stopwords(language = "en", source = "snowball")) 
tokens_NYT <- tokens_remove(tokens_NYT, c("s", "said", "would", "also", "one")) # additional stopwords

# tokens_NYT <- tokens_remove(tokens_NYT, c("said", "would")) # additional stoword list extended
# DFM 
dfm_NYT <- dfm(tokens_NYT)
dfm_NYT <- dfm_trim(dfm_NYT, min_termfreq = 10) # Trim rare words

# Convert to STM format
stm_input_NYT <- convert(dfm_NYT, to = "stm")
documents_NYT <- stm_input_NYT$documents
vocab_NYT <- stm_input_NYT$vocab
meta_NYT <- NYT_Input[match(docnames(dfm_NYT), NYT_Input$doc_id), ]




# NYT // defined metadata / covariates for the STM Input
meta_NYT$date_numeric <- as.numeric(as.Date(meta_NYT$date))
meta_NYT$country_iso <- ifelse(is.na(meta_NYT$country_iso), "none", as.character(meta_NYT$country_iso))
meta_NYT$country_n <- ifelse(is.na(meta_NYT$country_n), "none", as.character(meta_NYT$country_n))

meta_NYT$country_n <- as.factor(meta_NYT$country_n)
meta_NYT$country_iso <- as.factor(meta_NYT$country_iso)
```

## SearchK

SearchK is applied as an important component to identify the optimal number of topics *k* as discussed in section 3 Methods in the Masterthesis. This is applied on the Model 1 without covariates for both TS and NYT. Results indicate *k*=15 topics is a optimal tradeoff.

```{r}
#| label: SearchK
# # 5. searchK to find optimal number k of topics for stm with TS data
# k_result_TS <- searchK(
#   documents = documents_TS,
#   vocab = vocab_TS,
#   K = c(5, 10, 15, 20, 25, 30),
#   prevalence = ~ 1,
#   data = meta_TS,
#   init.type = "Spectral")
# 
# plot(k_result_TS)
# title(main = "Tagesschau Data",
#       line = 3,
#       cex.main = 1)
# 
# 
# 
# 
# # # 5. searchK to find optimal number k of topics for stm with NYT data
#  k_result_NYT <- searchK(
#    documents = documents_NYT,
#    vocab = vocab_NYT,
#    K = c(5, 10, 15, 20, 25, 30),
#    prevalence = ~ 1,
#    data = meta_NYT,
#    init.type = "Spectral")
# 
#  plot(k_result_NYT)
#  title(main = "NYT Data",
#        line = 3,
#        cex.main = 1)
```

## STM Model 1 (no covariates) for TS

This step runs the STM for TS (no covariates) and shows the output topwords, also the labeling and dominant theta is applied. Labeling decisions are discussed in section 3 Methods and Keyword_List.xlsx.

After the stm is run and labeled the output is aggregated on (1) daily level for descriptive analysis of topic distributions and (2) most importantly on weekly level to match the GT output. This allows to compare Media Attenton (TS) with Pulic Interest (GT) on the same weekly aggregated time series starting sunday each week.

The main output is "TS_Theta_Weekly_STM.csv" which is a weekly aggregated time series that will be compared to GT and be used for robustness checks.

```{r}
#| label: STM TS no Cov
# STM Model
stm_model_TS <- stm(
  documents = documents_TS,
  vocab = vocab_TS,
  K = 15, # 15 topics
  prevalence = ~ 1, # no use of date
  data = meta_TS,
  max.em.its = 75,
  init.type = "Spectral",
  seed = 2026)

# Topics Top Words (unlabeled)
labelTopics(stm_model_TS, n = 10)

theta_TS <- stm_model_TS$theta
theta_df_TS <- as.data.frame(theta_TS)
colnames(theta_df_TS) <- paste0("topic_", 1:ncol(theta_df_TS))
TS_theta <- bind_cols(meta_TS, theta_df_TS)

# Label Topics (manual based on Top Words)
topic_labels_TS <- c(
  "Natural Disasters",        # Topic 1 
  "EU Politics",              # Topic 2 
  "else",                     # Topic 3 (drop)
  "US Politics",              # Topic 4 
  "University & Protests",    # Topic 5
  "Military & Geopolitics",   # Topic 6
  "else",                     # Topic 7 (drop)
  "Ukraine War",              # Topic 8 
  "EU Right-wing Politics",   # Topic 9
  "Diplomacy",                # Topic 10
  "Israel Gaza Conflict",     # Topic 11 
  "Fires & Climate Disasters",# Topic 12
  "Accidents",                # Topic 13
  "Elections",                # Topic 14
  "else"                      # Topic 15 (drop)
)

# dominant topic index
dominant_topic_TS <- apply(theta_TS, 1, which.max)
dominant_theta_TS <- apply(theta_TS, 1, max) # dominant theta value
dominant_label_TS <- topic_labels_TS[dominant_topic_TS]

# add labels to TS data
TS_Output <- meta_TS %>%
  mutate(
    dominant_topic = dominant_topic_TS,
    dominant_theta = dominant_theta_TS,
    topic_label = dominant_label_TS)

table(TS_Output$topic_label)

# highlight all cases (TS articles) with dominant theta <40 (no clear topic assignment) as else
TS_Output <- TS_Output %>%
  mutate(
    topic_label_clean = ifelse(dominant_theta < 0.4, "else", topic_label))



# The following code aggregates the data on daily and weekly level to make them comparable to GT output.

# Create aggregated TS Dataset Daily
TS_theta$date <- as.Date(TS_theta$date) # Date format
theta_cols <- paste0("topic_", 1:15) # Topic-cols

# Daily Aggregation (all topics)
TS_Theta_Daily <- TS_theta %>%
  group_by(date) %>%
  summarise(across(all_of(theta_cols), \(x) sum(x, na.rm = TRUE)),
            .groups = "drop")

# clean variable names: remove spaces & special characters completely
clean_names <- gsub("[^A-Za-z0-9]", "", topic_labels_TS)
clean_names <- make.unique(clean_names)
colnames(TS_Theta_Daily)[-1] <- clean_names


# Create aggregated TS Dataset Weekly
TS_Theta_Daily$date <- as.Date(TS_Theta_Daily$date)
theta_cols <- names(TS_Theta_Daily)[names(TS_Theta_Daily) != "date"]

# function: round to last sunday (GT logic to match GT Data)
get_sunday <- function(x) {
  x - wday(x, week_start = 1) %% 7}

# Weekly Aggregation
TS_Theta_Weekly <- TS_Theta_Daily %>%
  mutate(
    week_start = get_sunday(date),
    KW = isoweek(week_start),
    year = year(week_start),
    KW_Y = paste0("KW_", KW, "_", year)
  ) %>%
  group_by(week_start, KW, year, KW_Y) %>%
  summarise(
    across(all_of(theta_cols), \(x) sum(x, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  arrange(week_start)


# 1. "else"-Topics removed
TS_Theta_Weekly <- TS_Theta_Weekly %>% 
  select(-matches("else"))

# 2. _norm Suffix for scaling (0-100 scale)
TS_Theta_Weekly <- TS_Theta_Weekly %>%
  mutate(across(NaturalDisasters:Elections, 
                ~ (. / max(., na.rm = TRUE)) * 100, 
                .names = "{.col}_norm"))

# Long format
TS_Theta_Weekly_long <- TS_Theta_Weekly %>%
  pivot_longer(
    cols = -c(week_start, KW, year, KW_Y),
    names_to = "topic",
    values_to = "theta")

# Long format
TS_Theta_Weekly_long <- TS_Theta_Weekly_long %>%
  group_by(topic) %>%
  mutate(theta_scaled = theta / max(theta))
```

## STM Model 1 (no covariates) for NYT

This follows the exact same steps as the TS STM with identical settings, labeling and data aggregation on daily and weekly level.

The Output is "NYT_Theta_Weekly_STM.csv" which is a weekly aggregated time series that will be compared to GT and be used for robustness checks.

```{r}
#| label: STM NYT no cov

# STM Model
stm_model_NYT <- stm(
  documents = documents_NYT,
  vocab = vocab_NYT,
  K = 15,
  prevalence = ~ 1,
  data = meta_NYT,
  max.em.its = 75,
  init.type = "Spectral",
  seed = 2026)

# Topics (unlabeled)
labelTopics(stm_model_NYT, n = 10)



# Theta extraction
theta_NYT <- stm_model_NYT$theta
theta_df_NYT <- as.data.frame(theta_NYT)
colnames(theta_df_NYT) <- paste0("topic_", 1:ncol(theta_df_NYT))
NYT_theta <- bind_cols(meta_NYT, theta_df_NYT)

# Label Topics (manual based on Top Words)
topic_labels_NYT <- c(
  "else",                        # Topic 1 (drop)
  "Asia Geopolitics",            # Topic 2 
  "Ukraine War",                 # Topic 3 
  "else",                        # Topic 4 (drop)
  "Crime & Justice",             # Topic 5 
  "Accidents",                   # Topic 6 
  "Syria & Middle East Politics",# Topic 7
  "else",                        # Topic 8 (drop)
  "Global Leaders & Conflict",   # Topic 9
  "Economic & Political Crises", # Topic 10
  "else",                        # Topic 11 (drop)
  "Israel Gaza Conflict",        # Topic 12 
  "else",                        # Topic 13 (drop)
  "Religion & Culture",          # Topic 14 
  "else"                         # Topic 15 (drop)
)

# dominant topic index
dominant_topic_NYT <- apply(theta_NYT, 1, which.max)
dominant_theta_NYT <- apply(theta_NYT, 1, max) # dominant theta value
dominant_label_NYT <- topic_labels_NYT[dominant_topic_NYT]

# add labels to NYT data
NYT_Output <- meta_NYT %>%
  mutate(
    dominant_topic = dominant_topic_NYT,
    dominant_theta = dominant_theta_NYT,
    topic_label = dominant_label_NYT)

table(NYT_Output$topic_label)

# highlight all cases (NYT articles) with dominant theta <40 (no clear topic assignment) as else
NYT_Output <- NYT_Output %>%
  mutate(
    topic_label_clean = ifelse(dominant_theta < 0.4, "else", topic_label) # theta treshold
  )

# safe data
write.csv(TS_Output, "TS_Output.csv")
write.csv(NYT_Output, "NYT_Output.csv")


# The following code aggregates the data on daily and weekly level to make them comparable to GT output.

# Create aggregated NYT Dataset Daily
NYT_theta$date <- as.Date(NYT_theta$date) # Date format
theta_cols <- paste0("topic_", 1:15) # Topic-cols

# Daily Aggregation (all topics)
NYT_Theta_Daily <- NYT_theta %>%
  group_by(date) %>%
  summarise(across(all_of(theta_cols), \(x) sum(x, na.rm = TRUE)),
            .groups = "drop")

# clean variable names: remove spaces & special characters completely
clean_names <- gsub("[^A-Za-z0-9]", "", topic_labels_NYT)
clean_names <- make.unique(clean_names) # handle duplicates like "else"
colnames(NYT_Theta_Daily)[-1] <- clean_names


# Create aggregated NYT Dataset Weekly
NYT_Theta_Daily$date <- as.Date(NYT_Theta_Daily$date)
theta_cols <- names(NYT_Theta_Daily)[names(NYT_Theta_Daily) != "date"] 

# function: round to last sunday (GT logic to match GT Data)
get_sunday <- function(x) {
  x - wday(x, week_start = 1) %% 7}

# Weekly Aggregation
NYT_Theta_Weekly <- NYT_Theta_Daily %>%
  mutate(
    week_start = get_sunday(date),
    KW = isoweek(week_start),
    year = year(week_start),
    KW_Y = paste0("KW_", KW, "_", year)
  ) %>%
  group_by(week_start, KW, year, KW_Y) %>%
  summarise(
    across(all_of(theta_cols), \(x) sum(x, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  arrange(week_start)

# 1. "else"-Topics removed
NYT_Theta_Weekly <- NYT_Theta_Weekly %>% 
  select(-matches("else"))

# 2. _norm Suffix variables (0-100 scale)
NYT_Theta_Weekly <- NYT_Theta_Weekly %>%
  mutate(across(AsiaGeopolitics:ReligionCulture, 
                ~ (. / max(., na.rm = TRUE)) * 100, 
                .names = "{.col}_norm"))


# Create long format (same as TS)
NYT_Theta_Weekly_long <- NYT_Theta_Weekly %>%
  pivot_longer(
    cols = -c(week_start, KW, year, KW_Y),
    names_to = "topic",
    values_to = "theta")

# Normalize per topic (same logic as TS)
NYT_Theta_Weekly_long <- NYT_Theta_Weekly_long %>%
  group_by(topic) %>%
  mutate(theta_scaled = theta / max(theta)) %>%
  ungroup()


```

# Plot STM Output Time Series for TS and NYT

This shows an short summary of the time series for TS and NYT and safes the resulting time series as csv in the base format and long format. Base format will be used as time series against GT and long format for simple and intuitive plots. The data remains the same.

```{r}
#| label: Plot Output (short check)

# PLOT Weekly Change TS

ggplot(TS_Theta_Weekly_long, aes(x = week_start, y = theta_scaled, color = topic)) +
  geom_line() +
  theme_minimal()

# Plot per topic (facet)
ggplot(TS_Theta_Weekly_long, aes(x = week_start, y = theta_scaled)) +
  geom_line() +
  facet_wrap(~topic, scales = "free_y") +
  theme_minimal()

write.csv(TS_Theta_Weekly, "TS_Theta_Weekly_STM.csv")
write.csv(TS_Theta_Weekly_long, "TS_Theta_Weekly_long_STM.csv")



# PLOT Weekly Change NYT

ggplot(NYT_Theta_Weekly_long, aes(x = week_start, y = theta_scaled, color = topic)) +
  geom_line() +
  theme_minimal()

# Plot per topic (facet)
ggplot(NYT_Theta_Weekly_long, aes(x = week_start, y = theta_scaled)) +
  geom_line() +
  facet_wrap(~topic, scales = "free_y") +
  theme_minimal()

# Save output
write.csv(NYT_Theta_Weekly, "NYT_Theta_Weekly_STM")
write.csv(NYT_Theta_Weekly_long, "NYT_Theta_Weekly_long_STM.csv")
```

## Model 2 (with covariates) for TS

The second model will not be the main model as discussed in section 3 Methods. It will be used as a robustness check to see if metadata (covariates) influence topic aggregation and therefore potentially show different results when comparing Media Attention to Public Interest (GT).

The preprocessing of all Metadata is executed in the preprocessing steps above. Covariates include date, country_iso and country_n for each document.

By comparing the changes in theta distribution from Model 1 (no covariates) and Model 2 (covariates) on document level, it can be seen if the model adapts to metadata (Agreement Rate in %). Also the topwords may be influenced so topic order and labels could differ. This is discussed in the Masterthesis Results section on robustness checks. This approach follows the dominant theta approach and not the aggregated approach, to compare on document level.

```{r}
#| label: Model 2 TS with covariates

# STM Model with Covariates
stm_model_TS_cov <- stm(
  documents = documents_TS,
  vocab = vocab_TS,
  K = 15, # 15 topics
  prevalence = ~ s(date_numeric) + country_n + country_iso,
  data = meta_TS,
  max.em.its = 75,
  init.type = "Spectral",
  seed = 2026)

# Topics Top Words (unlabeled)
labelTopics(stm_model_TS_cov, n = 10)

theta_TS_cov <- stm_model_TS_cov$theta
theta_df_TS_cov <- as.data.frame(theta_TS_cov)

colnames(theta_df_TS_cov) <- paste0("topic_", 1:ncol(theta_df_TS_cov))

# Wichtig: meta_TS muss die gleiche Zeilenanzahl haben wie dfm_TS
TS_theta_cov <- bind_cols(meta_TS, theta_df_TS_cov)

# Manual Labels
topic_labels_TS_cov <- c(
  "Natural Disasters",           # Topic 1 
  "EU Politics",                 # Topic 2 
  "else",                        # Topic 3 (drop)
  "US Politics",                 # Topic 4 
  "University & Protests",       # Topic 5 
  "Military & Geopolitics",      # Topic 6 
  "else",                        # Topic 7 (drop)
  "Ukraine War",                 # Topic 8 
  "EU Right-wing Politics",      # Topic 9 
  "Diplomacy",                   # Topic 10
  "Israel Gaza Conflict",        # Topic 11
  "Fires & Climate Disasters",   # Topic 12
  "Accidents",                   # Topic 13
  "Elections",                   # Topic 14
  "else"                         # Topic 15 (drop)
)

# Aggregation Weekly 
TS_weekly_theta_cov <- TS_theta_cov %>%
  mutate(week = floor_date(as.Date(date), "week")) %>%
  group_by(week) %>%
  summarise(across(starts_with("topic_"), sum))


# dominant topic index
dominant_topic_TS_cov <- apply(theta_TS_cov, 1, which.max)
dominant_theta_TS_cov <- apply(theta_TS_cov, 1, max) # dominant theta value
dominant_label_TS_cov <- topic_labels_TS_cov[dominant_topic_TS_cov]

# add labels to TS data
TS_Output_cov <- meta_TS %>%
  mutate(
    dominant_topic = dominant_topic_TS_cov,
    dominant_theta = dominant_theta_TS_cov,
    topic_label = dominant_label_TS_cov)

table(TS_Output_cov$topic_label)

# highlight all cases (TS articles) with dominant theta <40 (no clear topic assignment) as else
TS_Output_cov <- TS_Output_cov %>%
  mutate(
    topic_label_clean = ifelse(dominant_theta < 0.4, "else", topic_label))

# safe data
write.csv(TS_Output_cov, "TS_Output_cov.csv")




# Additional (optional) TS
# This part compares the base model (without covariates) with the new model output including covariates to check how theta distributions change
prep_check_TS <- estimateEffect(1:15 ~ s(date_numeric) + country_iso, 
                             stmobj = stm_model_TS_cov, 
                             metadata = meta_TS)
# Check for an "Spiky" Topic 
summary(prep_check_TS, topics = 8)
comparison_table_TS <- table(Base = TS_Output$topic_label, 
                          Covariate = TS_Output_cov$topic_label)
agreement_TS <- sum(diag(comparison_table_TS)) / sum(comparison_table_TS)
print(paste0("Agreement Rate: ", round(agreement_TS * 100, 2), "%"))
divergence_rate_TS <- 100 - (agreement_TS * 100)
print(paste0("Divergence Rate: ", round(divergence_rate_TS, 2), "%"))





# Create Dataset to compare to GT (same approach as for Model 1)
# Daily Aggregation (to match the cleaning logic of Model 1)
TS_Theta_Daily_cov <- TS_theta_cov %>%
  group_by(date = as.Date(date)) %>%
  summarise(across(starts_with("topic_"), sum, na.rm = TRUE), .groups = "drop")

# Rename columns with cleaned topic labels (removing "else")
clean_names_cov <- gsub("[^A-Za-z0-9]", "", topic_labels_TS_cov)
clean_names_cov <- make.unique(clean_names_cov)
colnames(TS_Theta_Daily_cov)[-1] <- clean_names_cov

# Weekly Aggregation (Sunday-logic to match GT)
TS_Theta_Weekly_cov <- TS_Theta_Daily_cov %>%
  mutate(week_start = date - (wday(date, week_start = 1) %% 7)) %>%
  group_by(week_start) %>%
  summarise(across(NaturalDisasters:Elections, sum, na.rm = TRUE), .groups = "drop") %>%
  select(-matches("else")) # Remove else topics

# Create Normalized variables and add suffix _cov to ALL columns
# This ensures Model 2 is clearly distinguishable from Model 1
TS_Theta_Weekly_cov <- TS_Theta_Weekly_cov %>%
  mutate(across(NaturalDisasters:Elections, 
                ~ (. / max(., na.rm = TRUE)) * 100, 
                .names = "{.col}_norm")) %>%
  rename_with(~ paste0(., "_cov"), -week_start)

# Save result
write.csv(TS_Theta_Weekly_cov, "TS_Theta_Weekly_cov_STM.csv", row.names = FALSE)



```

## Model 2 (with covariates) for NYT

This follows the exact steps of the TS STM with Covarites (see above).

```{r}
#| label: NYT Model 2 with Covariates
# NYT // define metadata / covariates for the STM Input

meta_NYT$date_numeric <- as.numeric(as.Date(meta_NYT$date))
meta_NYT$country_iso <- ifelse(is.na(meta_NYT$country_iso), "none", as.character(meta_NYT$country_iso))
meta_NYT$country_n <- ifelse(is.na(meta_NYT$country_n), "none", as.character(meta_NYT$country_n))

meta_NYT$country_n <- as.factor(meta_NYT$country_n)
meta_NYT$country_iso <- as.factor(meta_NYT$country_iso)


# STM Model with Covariates
stm_model_NYT_cov <- stm(
  documents = documents_NYT,
  vocab = vocab_NYT,
  K = 15, # 15 topics
  prevalence = ~ s(date_numeric) + country_n + country_iso,
  data = meta_NYT,
  max.em.its = 75,
  init.type = "Spectral",
  seed = 2026)

# Topics Top Words (unlabeled)
labelTopics(stm_model_NYT_cov, n = 10)


# Theta (Themenwahrscheinlichkeiten pro Dokument) 
theta_NYT_cov <- stm_model_NYT_cov$theta
theta_df_NYT_cov <- as.data.frame(theta_NYT_cov)

colnames(theta_df_NYT_cov) <- paste0("topic_", 1:ncol(theta_df_NYT_cov))

# Theta-values inkl, metadata (date, Country etc.) 
NYT_theta_cov <- bind_cols(meta_NYT, theta_df_NYT_cov)

# Manual Labeling of NYT Data (cov) based on Keyword_List.xlsx
topic_labels_NYT_cov <- c(
  "Diplomacy",                    # Topic 1 
  "Asia Geopolitics",             # Topic 2 
  "Ukraine War",                  # Topic 3 
  "International Law & Courts",   # Topic 4 
  "Crime & Justice",              # Topic 5 
  "Accidents",                    # Topic 6 
  "Syria & Middle East Politics", # Topic 7 
  "else",                         # Topic 8 (drop)
  "Global Leaders & Conflict",    # Topic 9 
  "Economic & Political Crises",  # Topic 10 
  "else",                         # Topic 11 (drop)
  "Israel Gaza Conflict",         # Topic 12 
  "else",                         # Topic 13 (drop)
  "Religion & Culture",           # Topic 14 
  "Venezuela"                     # Topic 15 
)

# Aggregation Weekly 
NYT_weekly_theta_cov <- NYT_theta_cov %>%
  mutate(week = floor_date(as.Date(date), "week")) %>%
  group_by(week) %>%
  summarise(across(starts_with("topic_"), sum))


# dominant topic index
dominant_topic_NYT_cov <- apply(theta_NYT_cov, 1, which.max)
dominant_theta_NYT_cov <- apply(theta_NYT_cov, 1, max) # dominant theta value
dominant_label_NYT_cov <- topic_labels_NYT_cov[dominant_topic_NYT_cov]

# add labels to NYT data
NYT_Output_cov <- meta_NYT %>%
  mutate(
    dominant_topic = dominant_topic_NYT_cov,
    dominant_theta = dominant_theta_NYT_cov,
    topic_label = dominant_label_NYT_cov)

table(NYT_Output_cov$topic_label)

# highlight all cases (NYT articles) with dominant theta <40 (no clear topic assignment) as else
NYT_Output_cov <- NYT_Output_cov %>%
  mutate(
    topic_label_clean = ifelse(dominant_theta < 0.4, "else", topic_label) # theta treshold
  )

# safe data
write.csv(NYT_Output_cov, "NYT_Output_cov.csv")





#--
# Additional (optional) for NYT
# This part compares the base model (without covariates) with the new model output including covariates to check how theta distributions changes for each document. This follows the dominant theta approach and not the aggregated approach.
prep_check_NYT <- estimateEffect(1:15 ~ s(date_numeric) + country_iso, 
                                 stmobj = stm_model_NYT_cov, 
                                 metadata = meta_NYT)
# Check for an "Spiky" Topic 
summary(prep_check_NYT, topics = 3)
comparison_table_NYT <- table(Base = NYT_Output$topic_label, 
                              Covariate = NYT_Output_cov$topic_label)
agreement_NYT <- sum(diag(comparison_table_NYT)) / sum(comparison_table_NYT)
print(paste0("Agreement Rate: ", round(agreement_NYT * 100, 2), "%"))
divergence_rate_NYT <- 100 - (agreement_NYT * 100)
print(paste0("Divergence Rate: ", round(divergence_rate_NYT, 2), "%"))



# Create Dataset to compare to GT (same approach as for Model 1)
# Daily Aggregation
NYT_Theta_Daily_cov <- NYT_theta_cov %>%
  group_by(date = as.Date(date)) %>%
  summarise(across(starts_with("topic_"), sum, na.rm = TRUE), .groups = "drop")

# Rename columns with cleaned topic labels
clean_names_NYT_cov <- gsub("[^A-Za-z0-9]", "", topic_labels_NYT_cov)
clean_names_NYT_cov <- make.unique(clean_names_NYT_cov)
colnames(NYT_Theta_Daily_cov)[-1] <- clean_names_NYT_cov

# Weekly Aggregation (Sunday-logic) and remove else
NYT_Theta_Weekly_cov <- NYT_Theta_Daily_cov %>%
  mutate(week_start = date - (wday(date, week_start = 1) %% 7)) %>%
  group_by(week_start) %>%
  summarise(across(Diplomacy:Venezuela, sum, na.rm = TRUE), .groups = "drop") %>%
  select(-matches("else"))

# Create Normalized variables and add suffix _cov to ALL columns
NYT_Theta_Weekly_cov <- NYT_Theta_Weekly_cov %>%
  mutate(across(Diplomacy:Venezuela, 
                ~ (. / max(., na.rm = TRUE)) * 100, 
                .names = "{.col}_norm")) %>%
  rename_with(~ paste0(., "_cov"), -week_start)

# Save result
write.csv(NYT_Theta_Weekly_cov, "NYT_Theta_Weekly_cov_STM.csv", row.names = FALSE)



```

## GT Data Input

This part adds the GT Data for each given Topics. The Processing ensures the same time series format as the TS and NYT output with weekly aggregated search volume per topic (normalized). The full process of GT keyword selection and queries is included in the Keywords_List.xlsx.

Note that the data is loaded from Keywords_List.xlsx , the file path needs to be changed accordingly.

```{r}
#| label: GT Data Loading and Processing
# File path to Keyword_List Excel (including all topics and related keywords)
file_path <- "/Keyword_List.xlsx"

# Get all sheet names (one topic per sheet, including keywors)
sheets <- excel_sheets(file_path)
keywords_list_df <- read_excel(file_path, sheet = "keywords_list")

# Remove metadata sheet (initial overview sheet explaining keyword finding process)
sheets <- sheets[!sheets %in% c("MT_Keywords_Table", "keywords_list")]

# Function: read, clean, aggregate duplicates (no duplicate dates and sheet names allowed)
read_and_process <- function(sheet_name) {
  
  GT_sheet <- read_excel(file_path, sheet = sheet_name)
  
  # Convert date column
  GT_sheet[[1]] <- as.Date(GT_sheet[[1]])
  
  # Convert keyword columns to numeric
  GT_sheet[-1] <- lapply(GT_sheet[-1], as.numeric)
  
  # Aggregate duplicate dates (important, otherwise duplicate dates)
  GT_sheet <- GT_sheet %>%
    group_by(across(1)) %>%
    summarise(across(everything(), ~ mean(.x, na.rm = TRUE)), .groups = "drop")
  
  # Clean keyword names (remove spaces)
  col_names <- colnames(GT_sheet)
  col_names[-1] <- gsub(" ", "", col_names[-1])
  
  # Rename columns: outlet_topic_keyword (so each column is easy to interpred and use for analysis)
  new_names <- c(
    "date",
    paste0(sheet_name, "_", col_names[-1])
  )
  
  colnames(GT_sheet) <- new_names
  
  return(GT_sheet)
}

# Read all sheets
GT_list <- lapply(sheets, read_and_process)

# Merge all sheets by date
GT_Data_raw <- reduce(GT_list, full_join, by = "date") %>%
  arrange(date)

#Build final dataset with correct column order

# Start with date (for loop to create GT df)
GT_Data <- GT_Data_raw %>% select(date)

for (topic in sheets) {
  
  # Select keyword columns for this topic
  topic_cols <- grep(paste0("^", topic, "_"), names(GT_Data_raw), value = TRUE)
  
  # Compute aggregates
  topic_sum <- rowSums(GT_Data_raw[topic_cols], na.rm = TRUE)
  topic_mean <- rowMeans(GT_Data_raw[topic_cols], na.rm = TRUE)
  
  # Build topic block: keywords + sum + mean
  topic_block <- GT_Data_raw %>%
    select(all_of(topic_cols)) %>%
    mutate(
      !!paste0(topic, "_sum") := topic_sum,
      !!paste0(topic, "_mean") := topic_mean
    )
  
  # Append to final dataset
  GT_Data <- bind_cols(GT_Data, topic_block)
}

# View result
head(GT_Data)

write.csv(GT_Data, "GT_Output_STM.csv")

```

## Merging Media Data (TS, NYT) and GT for Model 1

This step combines both time series (STM Model 1 and GT) into one dataframe. This will be the main dataset for the results analysis. Each topic from the media datasets is used in its normalized (\_norm) form and each corresponding GT result in its \_mean form. This allows to compare them efficiently as time series and plot them together.

```{r}
#| label: Merge Media (STM Model 1) and GT

# Merge Media Datasets (TS and NYT)
# This creates .x (TS) and .y (NYT) suffixes for duplicate column names
Model_1_Result_Input <- TS_Theta_Weekly %>%
  full_join(NYT_Theta_Weekly, by = c("week_start", "KW", "year", "KW_Y"))

# Join with GT Data
Model_1_Result_Input <- Model_1_Result_Input %>%
  left_join(GT_Data %>% rename(week_start = date), by = "week_start")

# Select and pair matching variables (using .x and .y for duplicates)
Model_1_Result_Input <- Model_1_Result_Input %>%
  select(
    week_start, KW, year, KW_Y,
    
    # TS Pairs: Media (topic_norm) and GT (topic_mean)
    TS_NaturalDisasters_norm = NaturalDisasters_norm, TS_NaturalDisasters_gt = TS_NaturalDisaster_mean,
    TS_EUPolitics_norm = EUPolitics_norm, TS_EUPolitics_gt = TS_EUPolitics_mean,
    TS_USPolitics_norm = USPolitics_norm, TS_USPolitics_gt = TS_USPolitics_mean,
    TS_MilitaryGeopolitics_norm = MilitaryGeopolitics_norm, TS_MilitaryGeopolitics_gt = TS_MilitaryGeopolitics_mean,
    TS_UkraineWar_norm = UkraineWar_norm.x, TS_UkraineWar_gt = TS_UkraineWar_mean,
    TS_EURightwing_norm = EURightwingPolitics_norm, TS_EURightwing_gt = TS_EURightWing_mean,
    TS_Diplomacy_norm = Diplomacy_norm, TS_Diplomacy_gt = TS_Diplomatie_mean,
    TS_IsraelGaza_norm = IsraelGazaConflict_norm.x, TS_IsraelGaza_gt = TS_IsraelGazaConflict_mean,
    TS_ClimateFires_norm = FiresClimateDisasters_norm, TS_ClimateFires_gt = TS_FiresClimateDisasters_mean,
    TS_Accidents_norm = Accidents_norm.x, TS_Accidents_gt = TS_Accidents_mean,
    TS_Elections_norm = Elections_norm, TS_Elections_gt = TS_Elections_mean,
    
    # NYT Pairs: Media (topic_norm) and GT (topic_mean)
    NYT_AsiaGeopolitics_norm = AsiaGeopolitics_norm, NYT_AsiaGeopolitics_gt = NYT_AsiaGeopolitics_mean,
    NYT_UkraineWar_norm = UkraineWar_norm.y, NYT_UkraineWar_gt = NYT_UkraineWar_mean,
    NYT_CrimeJustice_norm = CrimeJustice_norm, NYT_CrimeJustice_gt = NYT_CrimeJustice_mean,
    NYT_Accidents_norm = Accidents_norm.y, NYT_Accidents_gt = NYT_Accidents_mean,
    NYT_SyriaPolitics_norm = SyriaMiddleEastPolitics_norm, NYT_SyriaPolitics_gt = NYT_SyriaMiddleEastPolitics_mean,
    NYT_GlobalLeaders_norm = GlobalLeadersConflict_norm, NYT_GlobalLeaders_gt = NYT_GlobalLeaderConflict_mean,
    NYT_EconomicCrisis_norm = EconomicPoliticalCrises_norm, NYT_EconomicCrisis_gt = NYT_EconomicPoliticalCrisis_mean,
    NYT_IsraelGaza_norm = IsraelGazaConflict_norm.y, NYT_IsraelGaza_gt = NYT_IsraelGazaConflict_mean,
    NYT_ReligionCulture_norm = ReligionCulture_norm, NYT_ReligionCulture_gt = NYT_ReligionCulture_mean
  )

# Save final result basis
write.csv(Model_1_Result_Input, "Model_1_Result_Input.csv", row.names = FALSE)
```

## Merging Media Data (TS, NYT) and GT for Model 2

Same approach as for Model 1 but for the STM with Covariates including the \_cov suffix

```{r}
#| label: Merge Media (STM Model 2) and GT

# Merge Media Datasets (TS and NYT)
# This creates .x (TS) and .y (NYT) suffixes for all duplicate topic names
Model_2_Result_Input <- TS_Theta_Weekly_cov %>%
  full_join(NYT_Theta_Weekly_cov, by = "week_start")

# Join with GT Data
Model_2_Result_Input <- Model_2_Result_Input %>%
  left_join(GT_Data %>% rename(week_start = date), by = "week_start")

# Select and pair matching variables (handling .x/.y for all 4 overlaps)
Model_2_Result_Input <- Model_2_Result_Input %>%
  select(
    week_start,
    
    # TS Pairs: Media (topic_norm_cov) and GT (topic_mean)
    TS_NaturalDisasters_norm_cov = NaturalDisasters_norm_cov, TS_NaturalDisasters_gt = TS_NaturalDisaster_mean,
    TS_EUPolitics_norm_cov = EUPolitics_norm_cov, TS_EUPolitics_gt = TS_EUPolitics_mean,
    TS_USPolitics_norm_cov = USPolitics_norm_cov, TS_USPolitics_gt = TS_USPolitics_mean,
    TS_MilitaryGeopolitics_norm_cov = MilitaryGeopolitics_norm_cov, TS_MilitaryGeopolitics_gt = TS_MilitaryGeopolitics_mean,
    TS_UkraineWar_norm_cov = UkraineWar_norm_cov.x, TS_UkraineWar_gt = TS_UkraineWar_mean,
    TS_EURightwing_norm_cov = EURightwingPolitics_norm_cov, TS_EURightwing_gt = TS_EURightWing_mean,
    TS_Diplomacy_norm_cov = Diplomacy_norm_cov.x, TS_Diplomacy_gt = TS_Diplomatie_mean,
    TS_IsraelGaza_norm_cov = IsraelGazaConflict_norm_cov.x, TS_IsraelGaza_gt = TS_IsraelGazaConflict_mean,
    TS_ClimateFires_norm_cov = FiresClimateDisasters_norm_cov, TS_ClimateFires_gt = TS_FiresClimateDisasters_mean,
    TS_Accidents_norm_cov = Accidents_norm_cov.x, TS_Accidents_gt = TS_Accidents_mean,
    TS_Elections_norm_cov = Elections_norm_cov, TS_Elections_gt = TS_Elections_mean,
    
    # NYT Pairs: Media (topic_norm_cov) and GT (topic_mean)
    NYT_Diplomacy_norm_cov = Diplomacy_norm_cov.y, NYT_Diplomacy_gt = TS_Diplomatie_mean, # New overlap in M2
    NYT_AsiaGeopolitics_norm_cov = AsiaGeopolitics_norm_cov, NYT_AsiaGeopolitics_gt = NYT_AsiaGeopolitics_mean,
    NYT_UkraineWar_norm_cov = UkraineWar_norm_cov.y, NYT_UkraineWar_gt = NYT_UkraineWar_mean,
    NYT_IntLawCourts_norm_cov = InternationalLawCourts_norm_cov, NYT_IntLawCourts_gt = NYT_CrimeJustice_mean,
    NYT_CrimeJustice_norm_cov = CrimeJustice_norm_cov, NYT_CrimeJustice_gt = NYT_CrimeJustice_mean,
    NYT_Accidents_norm_cov = Accidents_norm_cov.y, NYT_Accidents_gt = NYT_Accidents_mean,
    NYT_SyriaPolitics_norm_cov = SyriaMiddleEastPolitics_norm_cov, NYT_SyriaPolitics_gt = NYT_SyriaMiddleEastPolitics_mean,
    NYT_GlobalLeaders_norm_cov = GlobalLeadersConflict_norm_cov, NYT_GlobalLeaders_gt = NYT_GlobalLeaderConflict_mean,
    NYT_EconomicCrisis_norm_cov = EconomicPoliticalCrises_norm_cov, NYT_EconomicCrisis_gt = NYT_EconomicPoliticalCrisis_mean,
    NYT_IsraelGaza_norm_cov = IsraelGazaConflict_norm_cov.y, NYT_IsraelGaza_gt = NYT_IsraelGazaConflict_mean,
    NYT_ReligionCulture_norm_cov = ReligionCulture_norm_cov, NYT_ReligionCulture_gt = NYT_ReligionCulture_mean,
    NYT_Venezuela_norm_cov = Venezuela_norm_cov, NYT_Venezuela_gt = NYT_GlobalLeaderConflict_mean
  )

# Save final result basis
write.csv(Model_2_Result_Input, "Model_2_Result_Input.csv", row.names = FALSE)
```
