---
title: "MT_04_Results"
author: "Fabian Gienke"
format: 
  html:
    toc: true           
    code-fold: true     
    code-summary: "Code anzeigen"
    theme: cosmo
execute:
  cache: true           
---

## Results Overview

The results are structured logically along the hypothesis. Plots and Tables are found in the section 4 Results and the Appendix in the Paper.

First descriptive Statistics about the Dataset are implemented. Then\
H1 (Correlations),\
H2 (Media Storms, Topic Categories),\
H3 (Time Lags, Granger),\
are analysed. Finally Robustness checks with Model 2 are implemented.

## R Packages

This chunk loads all required R packages for data preperation, time-series analysis, and formatting plots and tables.

```{r}
# Load all packages
library(tidyverse)
library(dplyr)
library(tidyr)
library(gt)
library(ggplot2)
library(lubridate)
library(vars)
library(lmtest)
library(tseries)
library(knitr)
library(xtable)

```

## Additional Data processing for Results

This section imports the weekly aggregated model results for both Model 1 and Model 2, alongside the article-level datasets containing the dominant theta distributions. File Paths need to be changed accordingly.

```{r}
#| label: Add Data

Model_1_Result_Input <- read.csv('/Model_1_Result_Input.csv')

Model_2_Result_Input <- read.csv('/Model_2_Result_Input.csv')

TS_Input_DominantTheta <- read.csv('/TS_Output.csv')

NYT_Input_DominantTheta <- read.csv('/NYT_Output.csv')
```

# Descriptive Statistics

## Overview Data: Media vs Public Table

This code calculates the average normalized media attention and public interest for each topic and formats the output into a descriptive summary table. Plot Table with Average Media Attention (norm) vs Average Public Interest (norm)

```{r}
#| label: Plot Table with Average Media Attention (norm) vs Average Public Interest (norm)
# 1. Summary for TS
ts_summary <- Model_1_Result_Input %>%
  summarise(across(starts_with("TS_") & contains("_norm"), mean, na.rm = TRUE),
            across(starts_with("TS_") & contains("_gt"), mean, na.rm = TRUE)) %>%
  pivot_longer(everything(), names_to = "key", values_to = "value") %>%
  mutate(outlet = "Tagesschau",
         topic = gsub("TS_|_norm|_gt", "", key),
         type = ifelse(grepl("_norm", key), "avg_norm_theta", "avg_gt_score")) %>%
  dplyr::select(-key) %>%
  pivot_wider(names_from = type, values_from = value)

# 2. Summary for NYT
nyt_summary <- Model_1_Result_Input %>%
  summarise(across(starts_with("NYT_") & contains("_norm"), mean, na.rm = TRUE),
            across(starts_with("NYT_") & contains("_gt"), mean, na.rm = TRUE)) %>%
  pivot_longer(everything(), names_to = "key", values_to = "value") %>%
  mutate(outlet = "NYT",
         topic = gsub("NYT_|_norm|_gt", "", key),
         type = ifelse(grepl("_norm", key), "avg_norm_theta", "avg_gt_score")) %>%
  dplyr::select(-key) %>%
  pivot_wider(names_from = type, values_from = value)

# 3. Combine and create HTML output
final_table_html <- bind_rows(ts_summary, nyt_summary) %>%
  arrange(outlet, desc(avg_norm_theta)) %>%
  gt() %>%
  tab_header(
    title = "Descriptive Statistics: Model 1",
    subtitle = "Comparison of Average Media Attention and Public Interest (GT)"
  ) %>%
  fmt_number(
    columns = c(avg_norm_theta, avg_gt_score),
    decimals = 2
  ) %>%
  cols_label(
    topic = "Topic",
    outlet = "Outlet",
    avg_norm_theta = "Avg. Media Attention (Norm)",
    avg_gt_score = "Avg. Public Interest (GT)"
  ) %>%
  tab_options(
    table.width = pct(100)
  )

# Display table
final_table_html
gtsave(final_table_html, "Descriptive_Stats_Model_1.html")


```

## Standard Deviation

This calculates the standard deviation for both the media and public datasets to measure and compare the overall volatility of each topic.

```{r}
#| label: SD Table
# SD Summary for TS
ts_sd_summary <- Model_1_Result_Input %>%
  summarise(across(starts_with("TS_") & contains("_norm"), sd, na.rm = TRUE),
            across(starts_with("TS_") & contains("_gt"), sd, na.rm = TRUE)) %>%
  pivot_longer(everything(), names_to = "key", values_to = "value") %>%
  mutate(outlet = "Tagesschau",
         topic = gsub("TS_|_norm|_gt", "", key),
         type = ifelse(grepl("_norm", key), "sd_norm_theta", "sd_gt_score")) %>%
  dplyr::select(-key) %>%
  pivot_wider(names_from = type, values_from = value)

# SD Summary for NYT
nyt_sd_summary <- Model_1_Result_Input %>%
  summarise(across(starts_with("NYT_") & contains("_norm"), sd, na.rm = TRUE),
            across(starts_with("NYT_") & contains("_gt"), sd, na.rm = TRUE)) %>%
  pivot_longer(everything(), names_to = "key", values_to = "value") %>%
  mutate(outlet = "NYT",
         topic = gsub("NYT_|_norm|_gt", "", key),
         type = ifelse(grepl("_norm", key), "sd_norm_theta", "sd_gt_score")) %>%
  dplyr::select(-key) %>%
  pivot_wider(names_from = type, values_from = value)

# Combine and create HTML output
sd_table_html <- bind_rows(ts_sd_summary, nyt_sd_summary) %>%
  arrange(outlet, desc(sd_norm_theta)) %>%
  gt() %>%
  tab_header(
    title = "Descriptive Statistics: Model 1 (Standard Deviation)",
    subtitle = "Volatility Comparison of Media Attention and Public Interest (GT)"
  ) %>%
  fmt_number(
    columns = c(sd_norm_theta, sd_gt_score),
    decimals = 2
  ) %>%
  cols_label(
    topic = "Topic",
    outlet = "Outlet",
    sd_norm_theta = "SD Media Attention",
    sd_gt_score = "SD Public Interest (GT)"
  ) %>%
  tab_options(
    table.width = pct(100)
  )

# Display and save
sd_table_html
gtsave(sd_table_html, "SD_Stats_Model_1.html")
```

## Summary Table for descriptive stats

This script aggregates article counts, weekly averages, and maximum search interest to create a comprehensive descriptive Latex table for the dataset.

```{r}
#| label: Summary Descriptive Stats Table


# compute metrics from weekly aggregated data
get_weekly_metrics <- function(data, outlet) {
  # identify media columns ending in _norm
  cols <- names(data)[grep(paste0("^", outlet, "_.*_norm$"), names(data))]
  
  map_dfr(cols, function(c) {
    topic <- gsub(paste0("^", outlet, "_|_norm$"), "", c)
    # identify corresponding google trends column
    gt_col <- paste0(outlet, "_", topic, "_gt")
    
    if(gt_col %in% names(data)) {
      tibble(
        Outlet = outlet,
        Topic = topic,
        `Avg Theta` = mean(data[[c]], na.rm = TRUE),
        `Avg GT` = mean(data[[gt_col]], na.rm = TRUE),
        `Max GT` = max(data[[gt_col]], na.rm = TRUE)
      )
    }
  })
}

# aggregate weekly stats for both outlets
weekly_summary <- bind_rows(
  get_weekly_metrics(Model_1_Result_Input, "TS"),
  get_weekly_metrics(Model_1_Result_Input, "NYT")
)

# combine document level data for article counts
all_docs <- bind_rows(
  TS_Input_DominantTheta %>% mutate(Outlet = "TS"),
  NYT_Input_DominantTheta %>% mutate(Outlet = "NYT")
)

# calculate total articles per outlet for share calculation
outlet_totals <- all_docs %>%
  group_by(Outlet) %>%
  summarise(total = n())

# calculate article counts and shares per topic
topic_counts <- all_docs %>%
  filter(topic_label_clean != "else") %>%
  mutate(
    # align names with weekly data structure
    raw_label = gsub("[^A-Za-z0-9]", "", topic_label_clean),
    Topic = case_when(
      raw_label == "EURightwingPolitics" ~ "EURightwing",
      raw_label == "IsraelGazaConflict" ~ "IsraelGaza",
      raw_label == "FiresClimateDisasters" ~ "ClimateFires",
      raw_label == "SyriaMiddleEastPolitics" ~ "SyriaPolitics",
      raw_label == "GlobalLeadersConflict" ~ "GlobalLeaders",
      raw_label == "EconomicPoliticalCrises" ~ "EconomicCrisis",
      TRUE ~ raw_label
    )
  ) %>%
  group_by(Outlet, Topic) %>%
  summarise(n = n(), .groups = "drop") %>%
  left_join(outlet_totals, by = "Outlet") %>%
  mutate(share = (n / total) * 100) %>%
  dplyr::select(Outlet, Topic, `N Articles` = n, `% Articles` = share)

# join counts with weekly averages and max values
final_desc_table <- topic_counts %>%
  inner_join(weekly_summary, by = c("Outlet", "Topic")) %>%
  arrange(Outlet, desc(`N Articles`))

# generate latex table output
kable_table <- kable(final_desc_table, format = "latex", booktabs = TRUE, 
      digits = c(0, 0, 0, 1, 2, 2, 2),
      col.names = c("Outlet", "Topic", "N Articles", "% Articles", "Avg $\\theta$ (norm.)", "Avg GT", "Max GT"))

# print result to console
cat(kable_table, sep = "\n")
```

## Dominant Theta vs. Aggregated Approach

This compares the sheer number of articles assigned to each topic against their aggregated theta probability sum to validate the structural topic classification.

```{r}
#| label: dominant theta vs aggregated theta

# combine document level datasets for both outlets
all_docs <- bind_rows(
  TS_Input_DominantTheta %>% mutate(Outlet = "TS"),
  NYT_Input_DominantTheta %>% mutate(Outlet = "NYT")
)

# calculate total articles and total theta sum per outlet
outlet_totals <- all_docs %>%
  group_by(Outlet) %>%
  summarise(
    total_n = n(),
    total_theta = sum(dominant_theta, na.rm = TRUE),
    .groups = "drop"
  )

# calculate article counts and theta sums per topic
final_desc_table <- all_docs %>%
  filter(topic_label_clean != "else") %>%
  mutate(
    # clean topic labels to match papers naming convention
    raw_label = gsub("[^A-Za-z0-9]", "", topic_label_clean),
    Topic = case_when(
      raw_label == "EURightwingPolitics" ~ "EURightwing",
      raw_label == "IsraelGazaConflict" ~ "IsraelGaza",
      raw_label == "FiresClimateDisasters" ~ "ClimateFires",
      raw_label == "SyriaMiddleEastPolitics" ~ "SyriaPolitics",
      raw_label == "GlobalLeadersConflict" ~ "GlobalLeaders",
      raw_label == "EconomicPoliticalCrises" ~ "EconomicCrisis",
      TRUE ~ raw_label
    )
  ) %>%
  group_by(Outlet, Topic) %>%
  summarise(
    n_articles = n(),
    theta_sum = sum(dominant_theta, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # merge with outlet totals to compute shares
  left_join(outlet_totals, by = "Outlet") %>%
  mutate(
    perc_articles = (n_articles / total_n) * 100,
    perc_theta = (theta_sum / total_theta) * 100
  ) %>%
  # select and rename columns for final output
  dplyr::select(Outlet, Topic, `N Articles` = n_articles, `% Articles` = perc_articles, 
                `Theta Sum` = theta_sum, `% Theta Share` = perc_theta) %>%
  arrange(Outlet, desc(`N Articles`))

# generate latex table
kable_table <- kable(final_desc_table, format = "latex", booktabs = TRUE, 
      digits = c(0, 0, 0, 1, 1, 1),
      col.names = c("Outlet", "Topic", "N Articles", "% Articles", "Theta Sum", "% Theta Share"))

# print latex code to console
cat(kable_table, sep = "\n")

```

## H1 - Media Storm Plot

This code detects media storms using the thresholds (entry $> 2\sigma$, exit $< 1\sigma$) and visualizes these storm phases with yellow highlights in the time series.

```{r}
#| label: Media Storms and Plot

# create object to call function
plot_storm_series <- function(data, outlet_prefix) {
  
  # Identify Cols
  media_cols <- names(data)[grep(paste0("^", outlet_prefix, "_.*_norm$"), names(data))]
  
  # Create Data in Long Format for Plot
  long_data <- data %>%
    dplyr::select(week_start, all_of(media_cols)) %>%
    pivot_longer(-week_start, names_to = "key", values_to = "Intensity") %>%
    mutate(
      Topic = gsub(paste0("^", outlet_prefix, "_|_norm$"), "", key),
      week_start = as.Date(week_start)
    ) %>%
    group_by(Topic) %>%
    mutate(
      # Define Tresholds
      avg_val = mean(Intensity, na.rm = TRUE),
      sd_val = sd(Intensity, na.rm = TRUE),
      thresh_2sd = avg_val + (2 * sd_val), # Start-Schwelle
      thresh_1sd = avg_val + (1 * sd_val)  # Ende-Schwelle
    )
  
  # Apply Media Storm Conditions
  long_data <- long_data %>%
    group_split() %>%
    map_dfr(function(df_topic) {
      in_storm <- logical(nrow(df_topic))
      current_state <- FALSE
      
      for(i in 1:nrow(df_topic)) {
        val <- df_topic$Intensity[i]
        if(!is.na(val)) {
          if(!current_state) {
            if(val > df_topic$thresh_2sd[i]) current_state <- TRUE
          } else {
            if(val < df_topic$thresh_1sd[i]) current_state <- FALSE
          }
        }
        in_storm[i] <- current_state
      }
      df_topic$is_storm <- in_storm
      return(df_topic)
    })
  
  # Red Lines for Tresholds and mean
  threshold_lines <- long_data %>%
    group_by(Topic) %>%
    summarise(
      Mean = first(avg_val),
      SD1 = first(thresh_1sd),
      SD2 = first(thresh_2sd),
      .groups = "drop"
    )
  
  # Highlight Storm Phases
  storm_highlights <- long_data %>%
    filter(is_storm == TRUE) %>%
    group_by(Topic, grp = cumsum(c(1, diff(week_start) > 7))) %>%
    summarise(xmin = min(week_start) - 3.5, xmax = max(week_start) + 3.5, .groups = "drop")
  
  # Plotting
  ggplot(long_data, aes(x = week_start, y = Intensity)) +
    geom_rect(data = storm_highlights, 
              aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = 100), 
              fill = "yellow", alpha = 0.3, inherit.aes = FALSE) +
    geom_hline(data = threshold_lines, aes(yintercept = Mean), color = "red", linetype = "dotted", alpha = 0.7) +
    geom_hline(data = threshold_lines, aes(yintercept = SD1), color = "red", linetype = "dashed", alpha = 0.7) +
    geom_hline(data = threshold_lines, aes(yintercept = SD2), color = "red", linetype = "solid", alpha = 0.8) +
    geom_line(color = "#003366", linewidth = 0.5) +
    facet_wrap(~Topic, ncol = 3, scales = "free_y") +
    scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 50)) +
    scale_x_date(date_labels = "%b %y", date_breaks = "6 months") +
    labs(
      title = paste("Thematic Dynamics & Media Storms:", outlet_prefix),
      subtitle = "Solid red: Mean+2SD (Entry) | Dashed: Mean+1SD (Exit) | Dotted: Mean",
      x = "Time (Weekly)",
      y = "Intensity (0-100)",
      caption = "Hysteresis detection: Start > 2SD, End < 1SD[cite: 1]"
    ) +
    theme_minimal(base_size = 9) +
    theme(
      strip.text = element_text(face = "bold"),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
}

# Create Plots
storm_plot_ts <- plot_storm_series(Model_1_Result_Input, "TS")
storm_plot_nyt <- plot_storm_series(Model_1_Result_Input, "NYT")

print(storm_plot_ts)
print(storm_plot_nyt)

```

## Output Table for Media Storms

This extracts the mathematically identified media storms and generates Latex tables detailing the total storm count, average duration, and maximum peak for each topic.

```{r}
#| label: Media Storm Table

# H1 - Media Storms

# function for plots
plot_storm_series <- function(data, outlet_prefix) {
  
  # identify cols
  media_cols <- names(data)[grep(paste0("^", outlet_prefix, "_.*_norm$"), names(data))]
  
  # create long data
  long_data <- data %>%
    dplyr::select(week_start, all_of(media_cols)) %>%
    pivot_longer(-week_start, names_to = "key", values_to = "Intensity") %>%
    mutate(
      Topic = gsub(paste0("^", outlet_prefix, "_|_norm$"), "", key),
      week_start = as.Date(week_start)
    ) %>%
    group_by(Topic) %>%
    mutate(
      avg_val = mean(Intensity, na.rm = TRUE),
      sd_val = sd(Intensity, na.rm = TRUE),
      thresh_2sd = avg_val + (2 * sd_val),
      thresh_1sd = avg_val + (1 * sd_val)
    )
  
  # apply identical storm logic
  long_data <- long_data %>%
    group_split() %>%
    map_dfr(function(df_topic) {
      
      in_storm <- logical(nrow(df_topic))
      current_state <- FALSE
      
      for(i in 1:nrow(df_topic)) {
        
        val <- df_topic$Intensity[i]
        
        if(!is.na(val)) {
          
          if(!current_state) {
            
            if(val > df_topic$thresh_2sd[i]) {
              current_state <- TRUE
            }
            
          } else {
            
            if(val < df_topic$thresh_1sd[i]) {
              current_state <- FALSE
            }
          }
        }
        
        in_storm[i] <- current_state
      }
      
      df_topic$is_storm <- in_storm
      
      return(df_topic)
    })
  
  # threshold lines
  threshold_lines <- long_data %>%
    group_by(Topic) %>%
    summarise(
      Mean = first(avg_val),
      SD1 = first(thresh_1sd),
      SD2 = first(thresh_2sd),
      .groups = "drop"
    )
  
  # highlight storm phases
  storm_highlights <- long_data %>%
    filter(is_storm == TRUE) %>%
    group_by(Topic, grp = cumsum(c(1, diff(week_start) > 7))) %>%
    summarise(
      xmin = min(week_start) - 3.5,
      xmax = max(week_start) + 3.5,
      .groups = "drop"
    )
  
  # plot
  ggplot(long_data, aes(x = week_start, y = Intensity)) +
    geom_rect(
      data = storm_highlights,
      aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = 100),
      fill = "yellow",
      alpha = 0.3,
      inherit.aes = FALSE
    ) +
    geom_hline(
      data = threshold_lines,
      aes(yintercept = Mean),
      color = "red",
      linetype = "dotted",
      alpha = 0.7
    ) +
    geom_hline(
      data = threshold_lines,
      aes(yintercept = SD1),
      color = "red",
      linetype = "dashed",
      alpha = 0.7
    ) +
    geom_hline(
      data = threshold_lines,
      aes(yintercept = SD2),
      color = "red",
      linetype = "solid",
      alpha = 0.8
    ) +
    geom_line(color = "#003366", linewidth = 0.5) +
    facet_wrap(~Topic, ncol = 3, scales = "free_y") +
    scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 50)) +
    scale_x_date(date_labels = "%b %y", date_breaks = "6 months") +
    labs(
      title = paste("Thematic Dynamics & Media Storms:", outlet_prefix),
      subtitle = "Solid red: Mean+2SD (Entry) | Dashed: Mean+1SD (Exit) | Dotted: Mean",
      x = "Time (Weekly)",
      y = "Intensity (0-100)",
      caption = "Hysteresis detection: Start > 2SD, End < 1SD"
    ) +
    theme_minimal(base_size = 9) +
    theme(
      strip.text = element_text(face = "bold"),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
}

# function for storm summary tables
create_storm_table <- function(data, outlet_prefix) {
  
  # identify cols
  media_cols <- names(data)[grep(paste0("^", outlet_prefix, "_.*_norm$"), names(data))]
  
  # create long data
  long_data <- data %>%
    dplyr::select(week_start, all_of(media_cols)) %>%
    pivot_longer(-week_start, names_to = "key", values_to = "Intensity") %>%
    mutate(
      Topic = gsub(paste0("^", outlet_prefix, "_|_norm$"), "", key),
      week_start = as.Date(week_start)
    ) %>%
    group_by(Topic) %>%
    mutate(
      avg_val = mean(Intensity, na.rm = TRUE),
      sd_val = sd(Intensity, na.rm = TRUE),
      thresh_2sd = avg_val + (2 * sd_val),
      thresh_1sd = avg_val + (1 * sd_val)
    )
  
  # apply identical storm logic
  long_data <- long_data %>%
    group_split() %>%
    map_dfr(function(df_topic) {
      
      in_storm <- logical(nrow(df_topic))
      current_state <- FALSE
      
      for(i in 1:nrow(df_topic)) {
        
        val <- df_topic$Intensity[i]
        
        if(!is.na(val)) {
          
          if(!current_state) {
            
            if(val > df_topic$thresh_2sd[i]) {
              current_state <- TRUE
            }
            
          } else {
            
            if(val < df_topic$thresh_1sd[i]) {
              current_state <- FALSE
            }
          }
        }
        
        in_storm[i] <- current_state
      }
      
      df_topic$is_storm <- in_storm
      
      return(df_topic)
    })
  
  # create storm ids
  storm_data <- long_data %>%
    group_by(Topic) %>%
    arrange(week_start) %>%
    mutate(
      storm_start = is_storm & !lag(is_storm, default = FALSE),
      storm_id = cumsum(storm_start)
    ) %>%
    ungroup()
  
  # calculate storm statistics
  storm_summary <- storm_data %>%
    filter(is_storm == TRUE) %>%
    group_by(Topic, storm_id) %>%
    summarise(
      duration_weeks = n(),
      peak_intensity = max(Intensity, na.rm = TRUE),
      peak_sd_above_mean = max((Intensity - avg_val) / sd_val, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    group_by(Topic) %>%
    summarise(
      number_storms = n(),
      average_duration = round(mean(duration_weeks), 2),
      maximum_peak_sd = round(max(peak_sd_above_mean), 2),
      .groups = "drop"
    ) %>%
    arrange(desc(number_storms))
  
  return(storm_summary)
}

# create plots
storm_plot_ts <- plot_storm_series(Model_1_Result_Input, "TS")
storm_plot_nyt <- plot_storm_series(Model_1_Result_Input, "NYT")

print(storm_plot_ts)
print(storm_plot_nyt)

# create tables
storm_table_ts <- create_storm_table(Model_1_Result_Input, "TS")
storm_table_nyt <- create_storm_table(Model_1_Result_Input, "NYT")

# latex tables
library(xtable)

xtable(
  storm_table_ts,
  caption = "Media Storm Characteristics for Tagesschau Topics",
  label = "tab:storm_ts"
)

xtable(
  storm_table_nyt,
  caption = "Media Storm Characteristics for New York Times Topics",
  label = "tab:storm_nyt"
)

# print latex code
print(
  xtable(
    storm_table_ts,
    caption = "Media Storm Characteristics for Tagesschau Topics",
    label = "tab:storm_ts"
  ),
  include.rownames = FALSE,
  booktabs = TRUE
)

print(
  xtable(
    storm_table_nyt,
    caption = "Media Storm Characteristics for New York Times Topics",
    label = "tab:storm_nyt"
  ),
  include.rownames = FALSE,
  booktabs = TRUE
)


```

## Aggregated Plot of Media Attention

This visualizes the overall corpus volatility by plotting the total weekly volume of labeled topic articles against the unclassified "else" category.

```{r}
#| label: Combined Time Series Plot


# create function to aggregate data weekly
get_corpus_volatility <- function(data, outlet_name) {
  
  # user specific sunday-week-start logic to match GT
  get_sunday <- function(x) { x - wday(x, week_start = 1) %% 7 }
  
  data %>%
    mutate(
      week_start = get_sunday(as.Date(date)),
      # categorize into labeled (signal) and else (noise)
      category = ifelse(topic_label_clean == "else", "Else Topics", "Labeled Topics")
    ) %>%
    group_by(week_start, category) %>%
    summarise(count = n(), .groups = "drop") %>%
    pivot_wider(names_from = category, values_from = count, values_fill = 0) %>%
    mutate(
      `All Topics (Total Volume)` = `Labeled Topics` + `Else Topics`,
      Outlet = outlet_name
    ) %>%
    # transform back to long format for ggplot
    pivot_longer(
      cols = c(`All Topics (Total Volume)`, `Labeled Topics`, `Else Topics`),
      names_to = "Series",
      values_to = "Volume"
    )
}

# apply function to both outlets
volatility_ts <- get_corpus_volatility(TS_Input_DominantTheta, "TS")
volatility_nyt <- get_corpus_volatility(NYT_Input_DominantTheta, "NYT")

# combine data
volatility_data <- bind_rows(volatility_ts, volatility_nyt)

# order factors so total volume is plotted distinctly in legend
volatility_data$Series <- factor(
  volatility_data$Series,
  levels = c("All Topics (Total Volume)", "Labeled Topics", "Else Topics")
)

# create professional plot
corpus_plot <- ggplot(volatility_data, aes(x = week_start, y = Volume, color = Series, linetype = Series)) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~Outlet, scales = "free_y", ncol = 2) +
  # updated color scheme: Green, Blue, Gray
  scale_color_manual(values = c(
    "All Topics (Total Volume)" = "#2E8B57", # Grün (SeaGreen)
    "Labeled Topics"            = "#003366", # Dunkelblau
    "Else Topics"               = "#999999"  # Grau
  )) +
  scale_linetype_manual(values = c(
    "All Topics (Total Volume)" = "dashed",
    "Labeled Topics" = "solid",
    "Else Topics" = "solid"
  )) +
  scale_y_continuous(limits = c(0, NA)) +
  scale_x_date(date_labels = "%b %y", date_breaks = "3 months") +
  labs(
    title = "Total Corpus Volatility: Signal vs. Noise",
    subtitle = "Weekly article volume of labeled topics versus unclassified (else) topics",
    x = "Time (Weekly)",
    y = "N Articles per Week",
    color = "Classification",
    linetype = "Classification"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold", size = 12),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# print and save plot
print(corpus_plot)
ggsave("Corpus_Volatility_Overview.pdf", corpus_plot, width = 20, height = 10, units = "cm")
```

## Combined Table of Media Attention Variance - full Corpus

This calculates summary statistics like the mean, standard deviation, and maximum peaks for the total weekly article volume across both outlets.

```{r}
# calculate summary statistics based on the existing volatility_data object
volatility_summary <- volatility_data %>%
  group_by(Outlet, Series) %>%
  summarise(
    `Total Articles` = sum(Volume, na.rm = TRUE),
    `Mean (Weekly)` = mean(Volume, na.rm = TRUE),
    `SD (Weekly)` = sd(Volume, na.rm = TRUE),
    `Max Peak` = max(Volume, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # sort by outlet and the predefined factor levels of Series
  arrange(Outlet, Series)

# generate latex table output
kable_volatility <- knitr::kable(
  volatility_summary, 
  format = "latex", 
  booktabs = TRUE, 
  digits = c(0, 0, 0, 1, 2, 0), # 1 decimal for mean, 2 for SD
  col.names = c("Outlet", "Classification", "Total Articles", "Mean (Weekly)", "SD", "Max Peak")
)

# print result to console
cat(kable_volatility, sep = "\n")
```

# H2 - Correlations

## Plot Time Series of Media and GT per topic

This function creates line plots to visually compare the normalized media coverage and Google Trends scores for specific individual topics. (Individual Time Series Plot)

```{r}
#| label: Plot per Topic of Time Series (Media and GT)
Model_1_Result_Input$week_start <- as.Date(Model_1_Result_Input$week_start)

plot_topic_comparison <- function(data, outlet, topic_name) {
  
  media_col <- paste0(outlet, "_", topic_name, "_norm")
  gt_col    <- paste0(outlet, "_", topic_name, "_gt")
  
  plot_data <- data %>%
    dplyr::select(week_start, all_of(media_col), all_of(gt_col)) %>%
    pivot_longer(cols = -week_start, names_to = "Series", values_to = "Score") %>%
    mutate(Series = ifelse(grepl("_norm", Series), "Media Attention", "Public Interest (GT)"))
  
  p <- ggplot(plot_data, aes(x = week_start, y = Score, color = Series)) +
    geom_line(linewidth = 0.8) +
    scale_color_manual(values = c("Media Attention" = "#003366", "Public Interest (GT)" = "#E69F00")) +
    scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 20)) +
    scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +
    labs(
      title = paste("Topic:", topic_name, "(", outlet, ")"),
      subtitle = "Comparison of Normalized Media Coverage and Google Search Volume",
      x = "Time (Weekly)",
      y = "Normalized Score (0-100)",
      color = ""
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom",
      plot.title = element_text(face = "bold", size = 14),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
  
  return(p)
}

p_ukraine <- plot_topic_comparison(Model_1_Result_Input, "TS", "UkraineWar")

# Display plot
print(p_ukraine)

ggsave("Plot_TS_UkraineWar.pdf", p_ukraine, width = 16, height = 9, units = "cm")

```

## Facet Plots

This creates large grid visualizations containing the time series of media attention and public interest for all topics simultaneously as facet plots.

```{r}


# Define function for faceted outlet overview
plot_outlet_grid <- function(data, outlet_prefix) {
  
  # Reshape data to long format for ggplot
  plot_data <- data %>%
    dplyr::select(week_start, starts_with(paste0(outlet_prefix, "_"))) %>%
    pivot_longer(-week_start, names_to = "key", values_to = "value") %>%
    mutate(
      # Extract clean topic names by removing prefix and suffix
      topic = gsub(paste0("^", outlet_prefix, "_|(_norm|_gt)$"), "", key),
      # Assign readable labels to series types
      type = ifelse(grepl("_norm", key), "Media Attention", "Public Interest (GT)")
    )

  ggplot(plot_data, aes(x = as.Date(week_start), y = value, color = type)) +
    geom_hline(yintercept = 0, color = "black", linewidth = 0.4) +
    geom_line(linewidth = 0.6) +
    facet_wrap(~topic, ncol = 3, scales = "fixed") + 
    scale_color_manual(values = c("Media Attention" = "#003366", "Public Interest (GT)" = "#E69F00")) +
    scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 50)) +
    scale_x_date(date_labels = "%y", date_breaks = "1 year") +
    labs(
      title = paste("Topic Comparison:", outlet_prefix),
      subtitle = "Time-series of Media Attention vs. Google Trends (2024-2025)",
      x = "Year",
      y = "Normalized Score (0-100)",
      color = ""
    ) +
    theme_minimal(base_size = 10) +
    theme(
      legend.position = "bottom",
      strip.text = element_text(face = "bold"),
      panel.spacing = unit(1, "lines"),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 0)
    )
}

# Generate grids for both outlets
grid_ts <- plot_outlet_grid(Model_1_Result_Input, "TS")
grid_nyt <- plot_outlet_grid(Model_1_Result_Input, "NYT")

# Save as PDF with A4-optimized dimensions
ggsave("Grid_TS_All_Topics.pdf", grid_ts, width = 18, height = 22, units = "cm")
ggsave("Grid_NYT_All_Topics.pdf", grid_nyt, width = 18, height = 20, units = "cm")
```

```{r}
# Safe as pdf
ggsave("Grid_TS_All_Topics.png", 
       grid_ts, 
       width = 18, 
       height = 22, 
       units = "cm", 
       dpi = 600,
       bg = "white")     
```

## Heatmap

This computes Pearson's correlation coefficients between media and public interest and visualizes the strength of these relationships as a color-coded heatmap.

```{r}
#| label: heatmap
# Calculate correlations between media and public interest per topic
calculate_correlations <- function(data, outlet_prefix) {
  
  # Identify all unique topics for the given outlet
  cols <- names(data)[grep(paste0("^", outlet_prefix, "_.*_norm$"), names(data))]
  topics <- gsub(paste0("^", outlet_prefix, "_|_norm$"), "", cols)
  
  # Loop through topics and compute Pearson correlation
  cor_results <- lapply(topics, function(t) {
    media_val <- data[[paste0(outlet_prefix, "_", t, "_norm")]]
    public_val <- data[[paste0(outlet_prefix, "_", t, "_gt")]]
    
    correlation <- cor(media_val, public_val, use = "complete.obs", method = "pearson")
    data.frame(outlet = outlet_prefix, topic = t, r = correlation)
  })
  
  bind_rows(cor_results)
}

# Combine results for both outlets
cor_data <- bind_rows(
  calculate_correlations(Model_1_Result_Input, "TS"),
  calculate_correlations(Model_1_Result_Input, "NYT")
)

# Create professional heatmap
heatmap_plot <- ggplot(cor_data, aes(x = outlet, y = reorder(topic, r), fill = r)) +
  geom_tile(color = "white", linewidth = 0.5) +
  # Use a professional blue-white-red or single hue gradient
  scale_fill_gradientn(
    colors = c("#f7fbff", "#6baed6", "#08306b"), 
    limits = c(0, 1),
    name = "Pearson's r"
  ) +
  # Add correlation coefficients as text inside tiles
  geom_text(aes(label = sprintf("%.2f", r)), color = ifelse(cor_data$r > 0.6, "white", "black"), size = 4) +
  labs(
    title = "Correlation: Media Attention vs. Public Interest",
    subtitle = "Pearson correlation coefficients per topic and outlet (Model 1)",
    x = "Media Outlet",
    y = "Topic",
    caption = "Source: STM Output & Google Trends Data"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    panel.grid = element_blank(),
    legend.position = "right"
  )

# Display and save as PDF
print(heatmap_plot)
ggsave("Correlation_Heatmap_Model1.pdf", heatmap_plot, width = 18, height = 18, units = "cm")
```

## Correlation with p values table

This script calculates the exact Pearson's correlations alongside their exact p-values, adding singificance stars to generate a formatted statistical table.

```{r}

# Function to calculate correlation and significance per topic
get_cor_stats <- function(data, outlet_prefix) {
  
  # Identify topics based on column naming convention
  cols <- names(data)[grep(paste0("^", outlet_prefix, "_.*_norm$"), names(data))]
  topics <- gsub(paste0("^", outlet_prefix, "_|_norm$"), "", cols)
  
  # Map results
  results <- lapply(topics, function(t) {
    m_val <- data[[paste0(outlet_prefix, "_", t, "_norm")]]
    p_val <- data[[paste0(outlet_prefix, "_", t, "_gt")]]
    
    # Perform Pearson correlation test
    test <- cor.test(m_val, p_val, method = "pearson", use = "complete.obs")
    
    # Extract values
    r_val <- test$estimate
    p_stat <- test$p.value
    
    # Define significance stars
    stars <- case_when(
      p_stat < 0.001 ~ "***",
      p_stat < 0.01  ~ "**",
      p_stat < 0.05  ~ "*",
      TRUE           ~ ""
    )
    
    data.frame(
      Outlet = outlet_prefix,
      Topic = t,
      r = r_val,
      p_value = p_stat,
      sig = stars,
      display_r = paste0(sprintf("%.2f", r_val), stars)
    )
  })
  
  bind_rows(results)
}

# Generate full stats table
cor_summary_table <- bind_rows(
  get_cor_stats(Model_1_Result_Input, "TS"),
  get_cor_stats(Model_1_Result_Input, "NYT")
) %>%
  arrange(desc(r))

# Create professional gt table for display/export
final_table_gt <- cor_summary_table %>%
  dplyr::select(Outlet, Topic, r, p_value, display_r) %>%
  gt() %>%
  tab_header(
    title = "Correlation Table: Media Attention vs. Public Interest",
    subtitle = "Statistical significance of thematic synchronicity (Pearson's r)"
  ) %>%
  cols_label(
    r = "Pearson's r",
    p_value = "p-value",
    display_r = "Correlation (sig.)"
  ) %>%
  fmt_number(columns = c(r, p_value), decimals = 3) %>%
  tab_source_note(
    source_note = "Significance levels: ***p < 0.001; **p < 0.01; *p < 0.05"
  ) %>%
  tab_options(
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )

# Show and save
final_table_gt
gtsave(final_table_gt, "Correlation_Stats_Model_1.html")
gtsave(final_table_gt, "Correlation_Stats_Model_1.pdf")
```

## Plot Single Topic Israel Gaza (TS)

This generates a single, high-contrast time-series plot specifically for the Israel & Gaza topic to serve as a detailed visual example in the main text.

```{r}
#| label: Plot Single Topic Israel Gaza (TS)

Model_1_Result_Input$week_start <- as.Date(Model_1_Result_Input$week_start)

outlet_name <- "TS"
topic_name  <- "IsraelGaza" 

media_col <- paste0(outlet_name, "_", topic_name, "_norm")
gt_col    <- paste0(outlet_name, "_", topic_name, "_gt")

plot_data_single <- Model_1_Result_Input %>%
  dplyr::select(week_start, all_of(media_col), all_of(gt_col)) %>%
  pivot_longer(cols = -week_start, names_to = "Series", values_to = "Score") %>%
  mutate(Series = ifelse(grepl("_norm", Series), "Media Attention", "Public Interest (GT)"))

p_israel_gaza <- ggplot(plot_data_single, aes(x = week_start, y = Score, color = Series)) +
  geom_line(linewidth = 0.8) +
  scale_color_manual(values = c("Media Attention" = "#003366", "Public Interest (GT)" = "#E69F00")) +
  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 20)) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +
  labs(
    title = "Time Series: Israel & Gaza (TS)",
    subtitle = "Comparison of Media Attention and Public Interest (2024-2025)",
    x = "Time (Weekly)",
    y = "Normalized Score (0-100)",
    color = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 13),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

print(p_israel_gaza)

ggsave("Plot_TS_IsraelGaza_Single.pdf", p_israel_gaza, width = 16, height = 9, units = "cm")
```

## Stationary ADF Test

Function to run Augmented Dickey-Fuller (ADF) tests for all topics in an outlet. This test confirms whether the underlying statistical properties are constant over time, which is a necessary condition for reliable Granger causality results and significance.

```{r}

# Function to run Augmented Dickey-Fuller (ADF) tests for all topics in an outlet
check_stationarity <- function(data, outlet_prefix) {
  media_cols <- names(data)[grep(paste0("^", outlet_prefix, "_.*_norm$"), names(data))]
  topics <- gsub(paste0("^", outlet_prefix, "_|_norm$"), "", media_cols)
  
  lapply(topics, function(t) {
    media_series <- na.omit(data[[paste0(outlet_prefix, "_", t, "_norm")]])
    public_series <- na.omit(data[[paste0(outlet_prefix, "_", t, "_gt")]])
    
    # Run ADF test (Null Hypothesis: Series is non-stationary)
    # A p-value < 0.05 indicates the series is stationary
    adf_media <- adf.test(media_series)
    adf_public <- adf.test(public_series)
    
    data.frame(
      Outlet = outlet_prefix,
      Topic = t,
      ADF_p_Media = adf_media$p.value,
      ADF_p_Public = adf_public$p.value,
      Stationary_Media = ifelse(adf_media$p.value < 0.05, "Yes", "No"),
      Stationary_Public = ifelse(adf_public$p.value < 0.05, "Yes", "No")
    )
  }) %>% bind_rows()
}

stationarity_results <- bind_rows(
  check_stationarity(Model_1_Result_Input, "TS"),
  check_stationarity(Model_1_Result_Input, "NYT")
)

print(stationarity_results)

non_stationary_cases <- stationarity_results %>% 
  filter(Stationary_Media == "No" | Stationary_Public == "No")

print(non_stationary_cases)
```

## First Differences Table

To ensure statistical validity, this part calculates Pearson's correlation using First Differences ($\Delta$) for any topic- Including information on the ADF tests stationarity.

```{r}
#| label: First Differences (All Topics)


# function to calculate first difference correlations for ALL topics
get_first_diff_correlations <- function(data, outlet_prefix) {
  
  cols <- names(data)[grep(paste0("^", outlet_prefix, "_.*_norm$"), names(data))]
  topics <- gsub(paste0("^", outlet_prefix, "_|_norm$"), "", cols)
  
  results <- lapply(topics, function(t) {
    media_raw <- data[[paste0(outlet_prefix, "_", t, "_norm")]]
    public_raw <- data[[paste0(outlet_prefix, "_", t, "_gt")]]
    
    valid_indices <- complete.cases(media_raw, public_raw)
    media <- media_raw[valid_indices]
    public <- public_raw[valid_indices]
    
    adf_media <- adf.test(media)$p.value
    adf_public <- adf.test(public)$p.value
    
    both_stationary <- ifelse((adf_media < 0.05) & (adf_public < 0.05), "Yes", "No")
    
    test <- cor.test(diff(media), diff(public), method = "pearson")
    
    # extract statistical values
    r_val <- test$estimate
    p_stat <- test$p.value
    
    # define significance stars
    stars <- case_when(
      p_stat < 0.001 ~ "***",
      p_stat < 0.01  ~ "**",
      p_stat < 0.05  ~ "*",
      TRUE           ~ ""
    )
    
    data.frame(
      Outlet = outlet_prefix,
      Topic = t,
      r = r_val,
      p_value = p_stat,
      display_r = paste0(sprintf("%.2f", r_val), stars),
      Stationary = both_stationary
    )
  })
  
  bind_rows(results)
}

first_diff_table <- bind_rows(
  get_first_diff_correlations(Model_1_Result_Input, "TS"),
  get_first_diff_correlations(Model_1_Result_Input, "NYT")
) %>%
  arrange(desc(r))

library(knitr)
first_diff_latex <- kable(
  first_diff_table %>% dplyr::select(Outlet, Topic, `Pearson's r (Δ)` = display_r, `p-value` = p_value, `Stationary (Raw)` = Stationary),
  format = "latex",
  booktabs = TRUE,
  digits = 3,
  row.names = FALSE,
  caption = "Correlation analysis of weekly changes (First Differences) between Media Attention and Public Interest. To ensure methodological consistency and avoid spurious correlations, all time series were transformed using First Differences. The 'Stationary (Raw)' column indicates whether the underlying raw time series were already stationary according to the ADF test.",
  label = "tab:correlations_first_diff"
)

cat(first_diff_latex, sep = "\n")
```

# H3 - Time Lags

## Basic Time Lags

This computes simple cross-correlations by shifting the time series up to two weeks forward and backward ($t-2$ to $t+2$) to identify basic lead-lag patterns.

```{r}

# Function to calculate lagged correlations and significance
calculate_lagged_cors <- function(data, outlet_prefix) {
  
  cols <- names(data)[grep(paste0("^", outlet_prefix, "_.*_norm$"), names(data))]
  topics <- gsub(paste0("^", outlet_prefix, "_|_norm$"), "", cols)
  
  results <- lapply(topics, function(t) {
    media <- data[[paste0(outlet_prefix, "_", t, "_norm")]]
    public <- data[[paste0(outlet_prefix, "_", t, "_gt")]]
    
    lags <- c(-2, -1, 0, 1, 2)
    
    lag_stats <- sapply(lags, function(l) {
      m_shifted <- if(l < 0) lag(media, abs(l)) else if(l > 0) lead(media, l) else media
      
      test <- cor.test(m_shifted, public, method = "pearson", use = "complete.obs")
      
      # Map p-values to significance stars
      stars <- case_when(
        test$p.value < 0.001 ~ "***",
        test$p.value < 0.01  ~ "**",
        test$p.value < 0.05  ~ "*",
        TRUE                 ~ ""
      )
      return(paste0(sprintf("%.2f", test$estimate), stars))
    })
    
    data.frame(
      Outlet = outlet_prefix,
      Topic = t,
      t_minus_2 = lag_stats[1],
      t_minus_1 = lag_stats[2],
      t_0 = lag_stats[3],
      t_plus_1 = lag_stats[4],
      t_plus_2 = lag_stats[5]
    )
  })
  
  bind_rows(results)
}

# Generate lag data for both outlets
lag_table_data <- bind_rows(
  calculate_lagged_cors(Model_1_Result_Input, "TS"),
  calculate_lagged_cors(Model_1_Result_Input, "NYT")
)

lag_table_final <- lag_table_data %>%
  gt(groupname_col = "Outlet") %>%
  tab_header(
    title = "Cross-Correlation Analysis with Time Lags",
    subtitle = "Pearson's r coefficients for Media vs. Public Interest (shifted by weeks)"
  ) %>%
  cols_label(
    Topic = "Topic",
    t_minus_2 = "t-2",
    t_minus_1 = "t-1",
    t_0 = "t (Simult.)",
    t_plus_1 = "t+1",
    t_plus_2 = "t+2"
  ) %>%
  tab_spanner(
    label = "Media Leads Public",
    columns = c(t_minus_2, t_minus_1)
  ) %>%
  tab_spanner(
    label = "Public Leads Media",
    columns = c(t_plus_1, t_plus_2)
  ) %>%
  tab_source_note(
    source_note = "Note: Media Leads indicates potential Agenda-Setting; Public Leads indicates potential Reverse Agenda-Setting. Stars: ***p < .001, **p < .01, *p < .05."
  ) %>%
  tab_options(
    table.font.size = 10,
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold"
  )

# View and save
lag_table_final
gtsave(lag_table_final, "Lag_Correlation_Table_M1.pdf")
```

## Granger Causality for stationary topics

This runs a standard Granger causality test at a 1-week lag exclusively for stationary topics, testing if media attention statistically predicts public interest or vice versa.

```{r}
#| label: Granger Analysis (Stationary Topics Only)


stationary_topics_ts <- c("IsraelGaza", "EURightwing", "Accidents", "UkraineWar", "Elections", "MilitaryGeopolitics", "NaturalDisasters")
stationary_topics_nyt <- c("IsraelGaza", "UkraineWar", "EconomicCrisis")

run_granger_stationary <- function(data, outlet_prefix, valid_topics) {
  
  results <- lapply(valid_topics, function(t) {
    media <- data[[paste0(outlet_prefix, "_", t, "_norm")]]
    public <- data[[paste0(outlet_prefix, "_", t, "_gt")]]
    
    # Granger Tests
    g_m_to_p <- grangertest(public ~ media, order = 1)
    p_m_to_p <- g_m_to_p$`Pr(>F)`[2]
    
    g_p_to_m <- grangertest(media ~ public, order = 1)
    p_p_to_m <- g_p_to_m$`Pr(>F)`[2]
    
    get_stars <- function(p) {
      case_when(
        p < 0.001 ~ "***", 
        p < 0.01  ~ "**", 
        p < 0.05  ~ "*", 
        p < 0.1   ~ ".", 
        TRUE      ~ "ns"
      )}
    
    display_topic <- case_when(
      t == "IsraelGaza" ~ "Israel & Gaza",
      t == "EURightwing" ~ "EU Rightwing",
      t == "MilitaryGeopolitics" ~ "Military & Geopolitics",
      t == "NaturalDisasters" ~ "Natural Disasters",
      t == "UkraineWar" ~ "Ukraine War",
      t == "EconomicCrisis" ~ "Economic Crisis",
      TRUE ~ t
    )
    
    data.frame(
      Outlet = outlet_prefix,
      Topic = display_topic,
      f_val_m_to_p = g_m_to_p$F[2],
      p_val_m_to_p = p_m_to_p,
      sig_m_to_p = get_stars(p_m_to_p),
      f_val_p_to_m = g_p_to_m$F[2],
      p_val_p_to_m = p_p_to_m,
      sig_p_to_m = get_stars(p_p_to_m)
    )
  })
  
  bind_rows(results)
}

granger_results_stat <- bind_rows(
  run_granger_stationary(Model_1_Result_Input, "TS", stationary_topics_ts),
  run_granger_stationary(Model_1_Result_Input, "NYT", stationary_topics_nyt)
)

latex_df <- granger_results_stat %>%
  mutate(
    f_val_m_to_p = sprintf("%.2f", f_val_m_to_p),
    p_val_m_to_p = sprintf("%.3f", p_val_m_to_p),
    f_val_p_to_m = sprintf("%.2f", f_val_p_to_m),
    p_val_p_to_m = sprintf("%.3f", p_val_p_to_m)
  ) %>%
  dplyr::select(
    Outlet, Topic,
    `F (M->P)` = f_val_m_to_p,
    `p (M->P)` = p_val_m_to_p,
    `Sig. ` = sig_m_to_p,     
    `F (P->M)` = f_val_p_to_m,
    `p (P->M)` = p_val_p_to_m,
    `Sig.` = sig_p_to_m
  )

# Latex Table
granger_latex <- kable(
  latex_df,
  format = "latex",
  booktabs = TRUE,
  row.names = FALSE,
  align = c("l", "l", "r", "r", "l", "r", "r", "l"),
  caption = "Results of the Granger causality analysis for strictly stationary foreign news topics (Lag = 1 week). The analysis tests whether weekly variations in media coverage significantly precede public interest or vice versa. Significance levels: *** p < .001, ** p < .01, * p < .05, . p < .1, ns = not significant.",
  label = "tab:granger_causality_stationary"
)

cat(granger_latex, sep = "\n")
```

## Extended Time Lag Analysis (First Differences) Media -\> Public

This robustness analysis tests whether weekly changes in Media Attention precede weekly changes in Public Interest across longer lags ($t1$--$t5$). In contrast to the standard Granger model, all topics are estimated using First Differences ($\Delta$) to reduce the risk of spurious lagged relationships caused by non-stationary trends.

```{r}
#| label: granger_first_diff_all_topics

# Function to run Granger causality for all lags
run_granger_fd <- function(data, outlet_prefix, max_lag = 5) {
  
  media_cols <- names(data)[grep(
    paste0("^", outlet_prefix, "_.*_norm$"),
    names(data)
  )]
  
  topics <- gsub(
    paste0("^", outlet_prefix, "_|_norm$"),
    "",
    media_cols
  )
  
  results <- map_dfr(topics, function(topic) {
    
    media_col <- paste0(outlet_prefix, "_", topic, "_norm")
    public_col <- paste0(outlet_prefix, "_", topic, "_gt")
    
    # Skip topics without GT variable
    if (!public_col %in% names(data)) {
      return(NULL)
    }
    
    # Apply first differences
    media_diff <- diff(data[[media_col]])
    public_diff <- diff(data[[public_col]])
    
    df <- tibble(
      media = media_diff,
      public = public_diff
    ) %>%
      drop_na()
    
    map_dfr(1:max_lag, function(lag_k) {
      
      # Estimate VAR model
      model <- tryCatch(
        VAR(df, p = lag_k, type = "const"),
        error = function(e) NULL
      )
      
      if (is.null(model)) {
        return(NULL)
      }
      
      # Media -> Public
      granger_result <- tryCatch(
        causality(model, cause = "media")$Granger,
        error = function(e) NULL
      )
      
      if (is.null(granger_result)) {
        return(NULL)
      }
      
      tibble(
        Outlet = outlet_prefix,
        Topic = topic,
        Lag = lag_k,
        F_value = as.numeric(granger_result$statistic),
        p_value = as.numeric(granger_result$p.value)
      )
    })
  })
  
  return(results)
}

# Run for both outlets
granger_fd_all <- bind_rows(
  run_granger_fd(Model_1_Result_Input, "TS"),
  run_granger_fd(Model_1_Result_Input, "NYT")
)

# Add significance stars
granger_fd_all <- granger_fd_all %>%
  mutate(
    sig = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01  ~ "**",
      p_value < 0.05  ~ "*",
      p_value < 0.10  ~ ".",
      TRUE ~ ""
    ),
    
    Topic_Label = paste(Outlet, Topic)
  )

# Plot
ggplot(
  granger_fd_all,
  aes(
    x = factor(Lag),
    y = reorder(Topic_Label, F_value),
    fill = F_value
  )
) +
  geom_tile(color = "white", linewidth = 0.3) +
  
  geom_text(
    aes(label = sig),
    size = 4
  ) +
  
  scale_fill_gradient(
    low = "grey90",
    high = "#003366",
    name = "F-value"
  ) +
  
  labs(
    title = "Lead--Lag Dynamics: Media Leading Public",
    subtitle = "Granger causality using First Differences (t1-t5)",
    x = "Lag (Weeks)",
    y = "Topic"
  ) +
  
  theme_minimal(base_size = 11) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(face = "bold"),
    axis.text.y = element_text(size = 9),
    plot.title = element_text(face = "bold"),
    legend.position = "right"
  )
```

## Extended Lag Analysis: Public Leading Media

This analysis tests whether weekly changes in Public Interest precede weekly changes in Media Attention across longer lag structures ($t1$--$t5$). In contrast to the standard Granger model, all topics are estimated using First Differences ($\Delta$) to reduce the risk of spurious lagged relationships caused by non-stationary trends. The analysis therefore includes reverse agenda-setting (public leads media).

```{r}
#| label: granger first diff all topics_reverse


# Function to run reverse Granger causality for all lags
run_granger_fd_reverse <- function(data, outlet_prefix, max_lag = 5) {
  
  media_cols <- names(data)[grep(
    paste0("^", outlet_prefix, "_.*_norm$"),
    names(data)
  )]
  
  topics <- gsub(
    paste0("^", outlet_prefix, "_|_norm$"),
    "",
    media_cols
  )
  
  results <- map_dfr(topics, function(topic) {
    
    media_col <- paste0(outlet_prefix, "_", topic, "_norm")
    public_col <- paste0(outlet_prefix, "_", topic, "_gt")
    
    # Skip topics without GT variable
    if (!public_col %in% names(data)) {
      return(NULL)
    }
    
    # Apply first differences
    media_diff <- diff(data[[media_col]])
    public_diff <- diff(data[[public_col]])
    
    df <- tibble(
      media = media_diff,
      public = public_diff
    ) %>%
      drop_na()
    
    map_dfr(1:max_lag, function(lag_k) {
      
      # Estimate VAR model
      model <- tryCatch(
        VAR(df, p = lag_k, type = "const"),
        error = function(e) NULL
      )
      
      if (is.null(model)) {
        return(NULL)
      }
      
      # Public -> Media
      granger_result <- tryCatch(
        causality(model, cause = "public")$Granger,
        error = function(e) NULL
      )
      
      if (is.null(granger_result)) {
        return(NULL)
      }
      
      tibble(
        Outlet = outlet_prefix,
        Topic = topic,
        Lag = lag_k,
        F_value = as.numeric(granger_result$statistic),
        p_value = as.numeric(granger_result$p.value)
      )
    })
  })
  
  return(results)
}

# Run for both outlets
granger_fd_reverse <- bind_rows(
  run_granger_fd_reverse(Model_1_Result_Input, "TS"),
  run_granger_fd_reverse(Model_1_Result_Input, "NYT")
)

# Add significance stars
granger_fd_reverse <- granger_fd_reverse %>%
  mutate(
    sig = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01  ~ "**",
      p_value < 0.05  ~ "*",
      p_value < 0.10  ~ ".",
      TRUE ~ ""
    ),
    
    Topic_Label = paste(Outlet, Topic)
  )

# Plot
ggplot(
  granger_fd_reverse,
  aes(
    x = factor(Lag),
    y = reorder(Topic_Label, F_value),
    fill = F_value
  )
) +
  geom_tile(color = "white", linewidth = 0.3) +
  
  geom_text(
    aes(label = sig),
    size = 4
  ) +
  
  scale_fill_gradient(
    low = "grey90",
    high = "#003366",
    name = "F-value"
  ) +
  
  labs(
    title = "Lead--Lag Dynamics: Public Leading Media",
    subtitle = "Granger causality using First Differences (t1-t5)",
    x = "Lag (Weeks)",
    y = "Topic"
  ) +
  
  theme_minimal(base_size = 11) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(face = "bold"),
    axis.text.y = element_text(size = 9),
    plot.title = element_text(face = "bold"),
    legend.position = "right"
  )
```

## Identify optimal lag per topic

This calculates the AIC-optimal lag (from 1 to 4 weeks maximizing F value) for every topic and creates a visualization / plot. This is important as the basic Ganger Causality Framework is stationary, meaning that only a fixed lag time intervall can be tested, ie. 1 week without this being the real underlying lag necessarily.

```{r}


# Function to run optimized Granger analysis including significance
get_optimized_granger_data <- function(data, outlet_prefix) {
  cols <- names(data)[grep(paste0("^", outlet_prefix, "_.*_norm$"), names(data))]
  topics <- gsub(paste0("^", outlet_prefix, "_|_norm$"), "", cols)
  
  lapply(topics, function(t) {
    # Extract media and public interest series
    media <- data[[paste0(outlet_prefix, "_", t, "_norm")]]
    public <- data[[paste0(outlet_prefix, "_", t, "_gt")]]
    
    # Create a matrix and remove any missing values for statistical consistency
    temp_mat <- cbind(media, public) %>% na.omit()

    lag_sel <- VARselect(temp_mat, lag.max = 4, type = "both")
    opt_lag <- lag_sel$selection["AIC(n)"]
    
    # Run the Granger test at the mathematically optimal lag
    g_test <- grangertest(public ~ media, order = opt_lag)
    f_stat <- g_test$F[2]
    p_val <- g_test$`Pr(>F)`[2]
    
    # Categorize the results for visualization based on p-value thresholds
    # This ensures the plot distinguishes between 'noise' and 'valid leads'
    result_cat <- case_when(
      p_val < 0.05 ~ "Significant (p < .05)",
      p_val < 0.1  ~ "Marginal Trend (p < .1)",
      TRUE         ~ "Not Significant"
    )
    
    data.frame(
      Outlet = outlet_prefix,
      Topic = t,
      Optimal_Lag = opt_lag,
      F_Score = f_stat,
      P_Value = p_val,
      Result = result_cat
    )
  }) %>% bind_rows()
}

# Generate data for both Tagesschau and New York Times
granger_plot_data <- bind_rows(
  get_optimized_granger_data(Model_1_Result_Input, "TS"),
  get_optimized_granger_data(Model_1_Result_Input, "NYT")
)

# Create the visualization with significance-coded colors
ggplot(granger_plot_data, aes(x = reorder(Topic, Optimal_Lag), y = Optimal_Lag, fill = Result)) +
  geom_bar(stat = "identity", alpha = 0.9, width = 0.7) +
  coord_flip() +
  facet_wrap(~Outlet, scales = "free_y") +
  # Use blue tones for significant results and gray for non-significant ones
  scale_fill_manual(values = c(
    "Significant (p < .05)" = "#2c7bb6",
    "Marginal Trend (p < .1)" = "#abd9e9",
    "Not Significant" = "#d3d3d3"
  )) +
  scale_y_continuous(breaks = 1:4, limits = c(0, 4)) +
  labs(
    title = "Optimal Granger Lags & Statistical Significance",
    subtitle = "Bars indicate the AIC-selected lag; colors represent statistical validity",
    x = "News Topic",
    y = "Optimal Predictive Window (Weeks)",
    fill = "Test Result"
  ) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank())
```

## Facet Plots: Robustness Time Series of Media Attention vs. Google Trends (Model 2)

This re-plots the topic time-series grids using the alternative Model 2 dataset, visually demonstrating the smoothing impact of the added temporal covariates.

```{r}
#| label: facet-plots-model-2
#| fig-cap: "Time-series comparison of normalized Media Attention and Public Interest for Model 2."

plot_outlet_grid_m2 <- function(data, outlet_prefix) {
  
  # reshape data to long format for ggplot
  plot_data <- data %>%
    dplyr::select(week_start, starts_with(paste0(outlet_prefix, "_"))) %>%
    pivot_longer(-week_start, names_to = "key", values_to = "value") %>%
    mutate(
      # extract clean topic names by stripping prefix, _cov, _norm, and _gt 
      # this ensures that media data and gt data match perfectly to the same facet
      topic = gsub(paste0("^", outlet_prefix, "_|_cov|_norm|_gt"), "", key),
      
      # assign readable labels to series types
      type = ifelse(grepl("_norm", key), "Media Attention", "Public Interest (GT)")
    )

  # create the faceted time-series plot
  ggplot(plot_data, aes(x = as.Date(week_start), y = value, color = type)) +
    # baseline at zero
    geom_hline(yintercept = 0, color = "black", linewidth = 0.4) +
    geom_line(linewidth = 0.6) +
    facet_wrap(~topic, ncol = 3, scales = "fixed") + 
    scale_color_manual(values = c("Media Attention" = "#003366", "Public Interest (GT)" = "#E69F00")) +
    scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 50)) +
    scale_x_date(date_labels = "%y", date_breaks = "1 year") +
    labs(
      title = paste("Topic Comparison:", outlet_prefix, "(Model 2)"),
      subtitle = "Time-series of Media Attention vs. Google Trends (2024-2025)",
      x = "Year",
      y = "Normalized Score (0-100)",
      color = ""
    ) +
    theme_minimal(base_size = 10) +
    theme(
      legend.position = "bottom",
      strip.text = element_text(face = "bold"),
      panel.spacing = unit(1, "lines"),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 0)
    )
}

grid_ts_m2 <- plot_outlet_grid_m2(Model_2_Result_Input, "TS")
grid_nyt_m2 <- plot_outlet_grid_m2(Model_2_Result_Input, "NYT")

print(grid_ts_m2)
print(grid_nyt_m2)

# save as pdf with a4-optimized dimensions for the thesis appendix
ggsave("Grid_TS_All_Topics_M2.pdf", grid_ts_m2, width = 18, height = 22, units = "cm")
ggsave("Grid_NYT_All_Topics_M2.pdf", grid_nyt_m2, width = 18, height = 20, units = "cm")
```

## Robustness Full Correlation Table

This generates a hybrid correlation table for Model 2, applying regular Pearson's $r$ for stationary topics and First Differences for non-stationary ones to ensure valid robustness metrics.

```{r}
#| label: Hybrid Correlation Table Model 2
#| echo: false
#| results: asis

library(dplyr)
library(knitr)

stat_map <- data.frame(
  Outlet = c("NYT", "TS", "TS", "NYT", "TS", "TS", "TS", "TS", "TS", "TS", 
             "NYT", "NYT", "NYT", "TS", "NYT", "NYT", "TS", "NYT", "TS", "NYT"),
  Topic_Raw = c("SyriaPolitics", "IsraelGaza", "USPolitics", "IsraelGaza", "Accidents", 
                "EURightwing", "MilitaryGeopolitics", "NaturalDisasters", "ClimateFires", 
                "EUPolitics", "AsiaGeopolitics", "UkraineWar", "Accidents", "Elections", 
                "GlobalLeaders", "ReligionCulture", "Diplomacy", "CrimeJustice", 
                "UkraineWar", "EconomicCrisis"),
  Stationary = c("No", "Yes", "No", "Yes", "Yes", "Yes", "Yes", "Yes", "No", "No", 
                 "No", "Yes", "No", "Yes", "No", "No", "No", "No", "Yes", "Yes")
)

calc_hybrid_cor <- function(data, outlet, topic, is_stationary) {
  
  col_media <- paste0(outlet, "_", topic, "_norm_cov")
  col_public <- paste0(outlet, "_", topic, "_gt")
  
  if(!col_media %in% names(data) | !col_public %in% names(data)) return(NULL)
  
  media <- data[[col_media]]
  public <- data[[col_public]]
  
  valid_idx <- complete.cases(media, public)
  media <- media[valid_idx]
  public <- public[valid_idx]
  
  if(is_stationary == "Yes") {
    res <- cor.test(media, public, method = "pearson")
    method_label <- "Pearson's $r$"
  } else {
    res <- cor.test(diff(media), diff(public), method = "pearson")
    method_label <- "Pearson's $r(\\Delta)$"
  }
  
  p_val <- res$p.value
  est <- as.numeric(res$estimate) # Ensures a clean numeric value for sorting
  
  stars <- case_when(
    p_val < 0.001 ~ "***",
    p_val < 0.01  ~ "**",
    p_val < 0.05  ~ "*",
    TRUE          ~ ""
  )
  
  display_topic <- case_when(
    topic == "SyriaPolitics" ~ "Syria Politics",
    topic == "IsraelGaza" ~ "Israel \\& Gaza",
    topic == "USPolitics" ~ "US Politics",
    topic == "Accidents" ~ "Accidents",
    topic == "EURightwing" ~ "EU Rightwing",
    topic == "MilitaryGeopolitics" ~ "Military \\& Geopolitics",
    topic == "NaturalDisasters" ~ "Natural Disasters",
    topic == "ClimateFires" ~ "Climate \\& Fires",
    topic == "EUPolitics" ~ "EU Politics",
    topic == "AsiaGeopolitics" ~ "Asia Geopolitics",
    topic == "UkraineWar" ~ "Ukraine War",
    topic == "Elections" ~ "Elections",
    topic == "GlobalLeaders" ~ "Global Leaders",
    topic == "ReligionCulture" ~ "Religion \\& Culture",
    topic == "Diplomacy" ~ "Diplomacy",
    topic == "CrimeJustice" ~ "Crime \\& Justice",
    topic == "EconomicCrisis" ~ "Economic Crisis",
    TRUE ~ topic
  )
  
  data.frame(
    Outlet = paste0("\\ac{", outlet, "}"),
    Topic = display_topic,
    Method = method_label,
    Estimate_Num = est, # Used to safely sort
    Correlation = sprintf("%.3f%s", est, stars),
    P_Value = sprintf("%.3f", p_val),
    Stationary = is_stationary,
    stringsAsFactors = FALSE
  )
}

results_list <- list()
for(i in 1:nrow(stat_map)) {
  row_res <- calc_hybrid_cor(Model_2_Result_Input, stat_map$Outlet[i], stat_map$Topic_Raw[i], stat_map$Stationary[i])
  if(!is.null(row_res)) {
    results_list[[i]] <- row_res
  }
}

hybrid_df_m2 <- bind_rows(results_list) %>%
  arrange(desc(Estimate_Num)) %>%
  dplyr::select(
    Outlet, 
    Topic, 
    Method,
    Correlation, 
    `p-value` = P_Value, 
    `Stat. (ADF)` = Stationary
  )

latex_hybrid <- kable(
  hybrid_df_m2,
  format = "latex",
  booktabs = TRUE,
  row.names = FALSE,
  escape = FALSE, 
  align = c("l", "l", "l", "r", "r", "c"),
  caption = "Hybrid Correlation Analysis for Model 2. The table combines regular Pearson's $r$ for stationary topics with the First Differences approach $r(\\Delta)$ for non-stationary topics to ensure robust estimates across all dynamics. Significance levels: *** $p < 0.001$, ** $p < 0.01$, * $p < 0.05$.",
  label = "tab:results_corr_hybrid_m2"
)

cat(latex_hybrid, sep = "\n")
```

## H3 Granger Robustness t1 check with Model 2

```{r}
#| label: Hybrid Granger Causality Table Model 2
#| echo: false
#| results: asis

library(dplyr)
library(knitr)
library(lmtest)

stat_map <- data.frame(
  Outlet = c("NYT", "TS", "TS", "NYT", "TS", "TS", "TS", "TS", "TS", "TS", 
             "NYT", "NYT", "NYT", "TS", "NYT", "NYT", "TS", "NYT", "TS", "NYT"),
  Topic_Raw = c("SyriaPolitics", "IsraelGaza", "USPolitics", "IsraelGaza", "Accidents", 
                "EURightwing", "MilitaryGeopolitics", "NaturalDisasters", "ClimateFires", 
                "EUPolitics", "AsiaGeopolitics", "UkraineWar", "Accidents", "Elections", 
                "GlobalLeaders", "ReligionCulture", "Diplomacy", "CrimeJustice", 
                "UkraineWar", "EconomicCrisis"),
  Stationary = c("No", "Yes", "No", "Yes", "Yes", "Yes", "Yes", "Yes", "No", "No", 
                 "No", "Yes", "No", "Yes", "No", "No", "No", "No", "Yes", "Yes")
)

calc_hybrid_granger <- function(data, outlet, topic, is_stationary) {
  
  col_media <- paste0(outlet, "_", topic, "_norm_cov")
  col_public <- paste0(outlet, "_", topic, "_gt")
  
  if(!col_media %in% names(data) | !col_public %in% names(data)) return(NULL)
  
  media <- data[[col_media]]
  public <- data[[col_public]]
  
  valid_idx <- complete.cases(media, public)
  media <- media[valid_idx]
  public <- public[valid_idx]
  
  if(is_stationary == "Yes") {
    g_m_to_p <- grangertest(public ~ media, order = 1)
    g_p_to_m <- grangertest(media ~ public, order = 1)
    method_label <- "Granger (Raw)"
  } else {
    g_m_to_p <- grangertest(diff(public) ~ diff(media), order = 1)
    g_p_to_m <- grangertest(diff(media) ~ diff(public), order = 1)
    method_label <- "Granger ($\\Delta$)"
  }
  
  f_val_m2p <- g_m_to_p$F[2]
  p_val_m2p <- g_m_to_p$`Pr(>F)`[2]
  
  f_val_p2m <- g_p_to_m$F[2]
  p_val_p2m <- g_p_to_m$`Pr(>F)`[2]
  
  get_stars <- function(p) {
    case_when(
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      p < 0.1   ~ ".",
      TRUE      ~ "ns"
    )
  }
  
  display_topic <- case_when(
    topic == "SyriaPolitics" ~ "Syria Politics",
    topic == "IsraelGaza" ~ "Israel \\& Gaza",
    topic == "USPolitics" ~ "US Politics",
    topic == "Accidents" ~ "Accidents",
    topic == "EURightwing" ~ "EU Rightwing",
    topic == "MilitaryGeopolitics" ~ "Military \\& Geopolitics",
    topic == "NaturalDisasters" ~ "Natural Disasters",
    topic == "ClimateFires" ~ "Climate \\& Fires",
    topic == "EUPolitics" ~ "EU Politics",
    topic == "AsiaGeopolitics" ~ "Asia Geopolitics",
    topic == "UkraineWar" ~ "Ukraine War",
    topic == "Elections" ~ "Elections",
    topic == "GlobalLeaders" ~ "Global Leaders",
    topic == "ReligionCulture" ~ "Religion \\& Culture",
    topic == "Diplomacy" ~ "Diplomacy",
    topic == "CrimeJustice" ~ "Crime \\& Justice",
    topic == "EconomicCrisis" ~ "Economic Crisis",
    TRUE ~ topic
  )
  
  data.frame(
    Outlet = paste0("\\ac{", outlet, "}"),
    Topic = display_topic,
    Method = method_label,
    F_M2P = f_val_m2p,
    P_M2P = p_val_m2p,
    Sig_M2P = get_stars(p_val_m2p),
    F_P2M = f_val_p2m,
    P_P2M = p_val_p2m,
    Sig_P2M = get_stars(p_val_p2m),
    stringsAsFactors = FALSE
  )
}

granger_results_list <- list()
for(i in 1:nrow(stat_map)) {
  row_res <- calc_hybrid_granger(Model_2_Result_Input, stat_map$Outlet[i], stat_map$Topic_Raw[i], stat_map$Stationary[i])
  if(!is.null(row_res)) {
    granger_results_list[[i]] <- row_res
  }
}

hybrid_granger_m2 <- bind_rows(granger_results_list) %>%
  arrange(desc(F_M2P)) %>%
  mutate(
    F_M2P = sprintf("%.2f", F_M2P),
    P_M2P = sprintf("%.3f", P_M2P),
    F_P2M = sprintf("%.2f", F_P2M),
    P_P2M = sprintf("%.3f", P_P2M)
  ) %>%
  dplyr::select(
    Outlet, 
    Topic, 
    Method,
    `F (M->P)` = F_M2P,
    `p (M->P)` = P_M2P,
    `Sig. ` = Sig_M2P, 
    `F (P->M)` = F_P2M,
    `p (P->M)` = P_P2M,
    `Sig.` = Sig_P2M
  )

latex_granger_hybrid <- kable(
  hybrid_granger_m2,
  format = "latex",
  booktabs = TRUE,
  row.names = FALSE,
  escape = FALSE,
  align = c("l", "l", "l", "r", "r", "l", "r", "r", "l"),
  caption = "Hybrid Granger Causality Analysis for Model 2 (Lag = 1 week). The table applies standard Granger tests for stationary topics and First Differences $\\Delta$ for non-stationary topics to maintain econometric robustness. Significance levels: *** $p < .001$, ** $p < .01$, * $p < .05$, . $p < .1$, ns = not significant.",
  label = "tab:granger_hybrid_m2"
)

cat(latex_granger_hybrid, sep = "\n")
```
